### 2022-03-31 paper/paper.tex
Fleshing out technicals in paper, starting implementation

\documentclass[12pt,a4paper]{article}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{booktabs}
\usepackage[latin2]{inputenc}
\usepackage{t1enc}
\usepackage[mathscr]{eucal}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{graphics}
\numberwithin{equation}{section}
\usepackage{hyperref}
\usepackage[margin=2.9cm]{geometry}
\usepackage[outdir=./images/]{epstopdf}
\epstopdfsetup{outdir=./images/}
\usepackage{algorithm}
\usepackage{algpseudocode}

\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}

\def\numset#1{{\\mathbb #1}}

\begin{document}

\title{Estimating Gaze Duration Error Distributions from Eye Tracking Data}

\author{
John Hawkins \\ email \href{mailto:john.hawkins@playgroundxyz.com}{john.hawkins@playgroundxyz.com} \\
}

\maketitle

\section{Abstract}

Eye tracking applications produce a series of gaze fixation points that can be attributed to
objects within a subjects field of vision. Error is typically measured on the basis of individual
gaze fixation measurements. However, often these applications are used to infer a gaze duration
metric from a series of fixation measurements. In this work we deliver an open source application
that estimates the error distribution of gaze duration estimates from an eye tracking validation
log file.

\section{Introduction}

Eye tracking applications are an important technology that permit the development of a wide 
variety of new forms of software interaction, as well as a tool for empirical investigations 
into human thought and behaviour. Eye tracking studies are widely used as a method of 
measuring overt attention to visual stimuli and have been applied to study the extent to 
which people look at, and remember, advertising \cite{Hervet2011}.
The technology has also allowed the study of many factors that contribute to effective advertising,
including the impact of images of faces \cite{Djamasbi2010},
the use of animation \cite{Hamborg2012} and the relationship with
social media posts \cite{Barreto2013}.

Increasingly, eye tracking applications are built using machine learning models trained on 
data sets containing combinations of facial images and fixation coordinates collected 
through careful application design. The resulting gaze fixation prediction model can 
then be applied sequentially to determine how long a subject fixated on a given object.
Thus, gaze duration is derived from aggregation of gaze fixation.

Gaze fixation models will typically have an error profile that is dependent on both 
the nature of the media being presented, the variety of faces and lighting conditions 
in the training data. In addition, it has been observed that the error profile varies 
depending on the location of the true fixation point.

When fixation models are used sequentially for the purpose of estimating gaze duration 
(as is done in advertising media studies), then there is the potential for the error 
to either cancel out (reducing error in gaze duration estimation) or to compound 
(increasing error in gaze duration). Which of these outcomes occurs will depend on the 
specific of the model and the conditions under which it is being used.
Influencing factors include the ratio of true positives to true negatives in the data, 
and the error profile of the machine learning fixation coordinate model. 
Heretofore, gaze duration models have assumed the former situation, if indeed the possibility of error in gaze duration is even considered.

In this work we develop a methodology for estimating session specific error profiles that relies on an empirical error profile of the underlying fixation model. The method allows us to profile whether an individual session has a compounding or cancelling error profile, and put empirical confidence bounds on the gaze duration estimates.

\section{Methodology}

We estimate the gaze duration measurement error through Monte Carlo simulation
on the basis of the gaze fixation errors on the validation data from eye tracking
fixation model. This error profile can be stratified into a grid across the screen dimensions. For each of these grid locations we calculate two error probabilities, the false negative probability (when true fixation is on the grid and the model produces coordinates outside the grid. And the false positive probability, when the true fixation is outside the grid location and the model produces a prediction inside the grid square.

We apply this grid of error probabilities in a Monte Carlo simulation that requires the following inputs:
1.	The grid coordinates in which the item of concern is located.
2.	The duration of true fixation on that item
3.	The duration of the entire session of measurement.
The Monte Carlo process will then repeatedly simulate sessions of the given length and gaze duration. Each simulation will use the error profiles to produce a potential model output. Collectively these estimates model outputs constitute a probability distribution over outputs for the given scenario.

\subsection{Measurement Distribution}

In the first stage of the methodology we provide an estimation of the distribution of 
measurements for a given true gaze duration. These distributions are produced at multiple locations
across the screen, so that we have a set of distributions, that captures the expected
variation in measurement depending on the length of and locus of true fixation.

We want to estimate a the probability distribution over measured durations $\hat{d}$ given a
known duration $d$, the session length $s$ and the ad location $l$. 
We express this distribution as $P(\hat{d}|d,l,s)$. The
procedure for making this estimation involves iteratively generating random samples
of gaze fixation paths $\mathcal{F}$ consistent with the parameters $d$, $s$ and $l$. 
For each point $f$ in $\mathcal{F}$ we determine a measurement point $\hat{f}$ 
by drawing samples from the eye tracking validation set. 
These points determine the $\delta_x$ and $\delta_y$ error in the measurement
of gaze fixation point $f$. We draw these coordinates from the validation file such the
the probability of the point being chosen is propotional to the distance of its target
$x$ and $y$ from the coordinates of the point $f$ in the generated path. The complete algorithm
for estimating $P(\hat{d}|d,l,s)$ is shown in Algorithm \ref{alg:p_of_dhat}

\begin{algorithm}
\caption{Estimation of $P(\hat{d}|d,l,s)$}\label{alg:p_of_dhat}
\begin{algorithmic}
  \Require $\hat{d},l,s,C$
  \Ensure $P(\hat{d}|d,l,s)$
  \State $N \gets 2000$
  \State $D \gets floor(s \times 10)$ \Comment{Durations sampled at 100ms intervals}
  \State $P \gets Array(D+1,D+1)$ \Comment{Result is a 2 dimensional array}
  \If{$\hat{d} > s$}
    \State error
  \EndIf
  \For{$d \in 0 to D$} \Comment{Iterate over all D+1 possible durations}
    \For{$n \in 1 to N$} \Comment{Collect N samples}
      \State $p \gets generatePath(d,l,s)$
      \State $d_e \gets generateMeasurement(p,C)$  
      \State $P[d,d_e] += 1/2000$  
    \EndFor
  \EndFor
\end{algorithmic}
\end{algorithm}

The bulk of the work is done by two functions inside the simulation loop. The first function
$generatePath(d,l,s)$ takes a true gaze duration, position of ad and viewing session length and
generates a path of fixation points compatible with those parameters. The second function 
$generateMeasurement(p,C)$, takes the path and the eye tracking calibration data and produces
an sample of measured attention time. Repeated application of this process over different 
values of $d$ allows us to populate a two dimensional array where the first dimension is the
true duration and the second the estimated. Note: that our probability distribution is discrete
because we restrict ourselves to sampling every 100ms point between zero and the session length.

\subsection{Gaze Duration Distribution}

In real world applications we will have the true session length but rely on the model for the 
estimate of gaze duration. Meaning that the probability distribution we want is the distribution 
over true gaze durations for a given measurement, rather than the inverse which is what we 
have produced in the previous section. We produce an estimate of $P(d|\hat{d},l,s)$
through an application of Bayes' rule, as shown in Equation \ref{eq:p_of_d}.

\begin{equation}
\label{eq:p_of_d}
P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ P(\hat{d},l,s)  }
\end{equation}

We use a uniform prior for $P(d)$, meaning in the absence of additional information all
gaze durations less than the session length are equally likely. As our distributions are
discrete estimations of an underlying continuous distribution the value of $P(d)$ is
equal to $1/(1+floor(10*s))$
 We can estimate the value
of the denominator $P(\hat{d},l,s)$ by iterating over all values of d and summing the product
of $ P(\hat{d},l,s|d) \dot P(d)$. This fully explicated form is shown in 
Equation \ref{eq:p_of_d_full}.

\begin{equation}
\label{eq:p_of_d_full}
P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ \sum_{\delta \in D} P(\hat{d},l,s|\delta) \dot P(\delta)  }
\end{equation}

\section{Results}

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}

### 2022-03-31 gazerr-runner.py

@@ -1,0 +1,10 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+ 
+"""Convenience wrapper for running gazerr directly from source tree."""
+ 
+from gazerr.cli import main
+ 
+if __name__ == '__main__':
+    main()
+

### 2022-03-31 gazerr/cli.py

@@ -21,46 +21,69 @@ def run_simulation(df):
 
 ####################################################################################
 def main():
-    parser = argparse.ArgumentParser(description='Run Error Estimation Experiment.')
+    desc = 'Estimate Gaze Duration Error Distribution from Eye Tracking Data'
+    parser = argparse.ArgumentParser(description=desc)
 
-    parser.add_argument('input',
-                       metavar='input',
+    parser.add_argument('calibration_data',
+                       metavar='calibration_data',
                        type=str,
-                       help='Path to the model calibration data.')
+                       help='Path to the model calibration data contains targets and gaze coords')
 
+    parser.add_argument('measurement',
+                       metavar='measurement',
+                       type=float,
+                       help='Measured gaze duration.')
+
+    parser.add_argument('session_length',
+                       metavar='session_length',
+                       type=float,
+                       help='Measured gaze duration.')
+
+    parser.add_argument('ad_position',
+                       metavar='ad_position',
+                       type=str,
+                       help='Position of the ad in the viewport: top | mid | bottom')
     parser.add_argument('output',
                        metavar='output',
                        type=str,
                        help='Path to output file location.')
-
-    parser.add_argument('-f', '--force', action="store_true", help="Force the output to overwrite existing files")
-
-    # Execute the parse_args() method
+ 
+    parser.add_argument('-f', '--force', action="store_true", help="Force the output overwrite.")
     args = parser.parse_args()
+    data = args.calibration_data
+    measure = args.measurement
+    session = args.session_length
+    position = args.ad_position
+    output = args.output
 
-    input_path = args.input
-    output_path = args.output
-
-    if not os.path.isfile(input_path):
+    if not os.path.isfile(data):
         print(" ERROR")
-        print(" The input file '%s' does not exist" % input_path)
+        print(" The input file '%s' does not exist" % data)
         sys.exit()
 
-    if os.path.isfile(output_path) and not args.force:
+    if os.path.isfile(output) and not args.force:
         print(" ERROR")
-        print(" The output file '%s' already exists."% output_path)
+        print(" The output file '%s' already exists."% output)
         print("    Use -f [--force] to force an overwrite over the file")
-        sys.exit()
 
-    df = load_calibration_data(input_path)
-
+    df = load_calibration_data(data)
     result = run_simulation(df)
+    result.to_csv(output, index=False, header=True)
 
-    result.to_csv(output_path, index=False, header=True)
[…]

### 2022-03-31 scripts/analyse_log_data.py

@@ -1,0 +1,121 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import math
+
+df = pd.read_csv("data/eye_tracking_gaze_log.csv")
+print(f"Total records {len(df)}")
+
+df = df[df['eye_tracking_state']=='success']
+print(f"Successful records {len(df)}")
+print(f"From {min(df['timestamp'])} to {max(df['timestamp'])} ")
+
+print(f"Calibration Records {len(df[df['event_type']=='calibration'])}")
+print(f"Validation Records {len(df[df['event_type']=='validation'])}")
+
+print("Target Numbers for Calibration")
+print(df[df['event_type']=='calibration']['target_number'].value_counts())
+print("Target Numbers for Validation")
+print(df[df['event_type']=='validation']['target_number'].value_counts())
+
+def euclidean_distance(row):
+    xdiff = row['target_x'] - row['gaze_x']
+    ydiff = row['target_y'] - row['gaze_y']
+    return math.sqrt( math.pow(xdiff,2) + math.pow(ydiff,2) )
+
+df['ERR'] = df.apply(euclidean_distance,axis=1)
+
+# Basics
+print("Total mean:",df['ERR'].mean())
+print("Total median:",df['ERR'].median())
+print("calibration mean", df[df['event_type']=="calibration"]['ERR'].mean())
+print("calibration median", df[df['event_type']=="calibration"]['ERR'].median())
+print("validation mean", df[df['event_type']=="validation"]['ERR'].mean())
+print("validation median", df[df['event_type']=="validation"]['ERR'].median())
+
+dataset = df[df['event_type']=="calibration"]['target_number'].value_counts().reset_index()
+plt.scatter(dataset['index'], dataset['target_number'])
+plt.title("Record count by target_number for calibration data")
+plt.xlabel("target_number")
+plt.ylabel("count")
+plt.savefig("results/records_by_target_number.png")
+
+
+df['ERR'].hist(bins=100)
+plt.title("Error Distribution")
+plt.savefig("results/Error_Distribution.png",format='png', dpi=100, bbox_inches='tight')
+
+tech = ['eye_tracking_calibration_mode','device_model', 'os', 'os_version']
+groups = [ 'study_id', 'user_id', 'user_ext', 'session_id']
+
+# Create subplots 
[…]

### 2022-03-31 scripts/calculate_errors.py

@@ -1,0 +1,72 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import math
+
+df = pd.read_csv("data/eye_tracking_gaze_log.csv")
+
+df = df[df['eye_tracking_state']=='success']
+
+def euclidean_distance(row):
+    xdiff = row['target_x'] - row['gaze_x']
+    ydiff = row['target_y'] - row['gaze_y']
+    return math.sqrt( math.pow(xdiff,2) + math.pow(ydiff,2) )
+
+df['ERR'] = df.apply(euclidean_distance,axis=1)
+
+group_cols = ['user_ext', 'event_type', 'target_state', 'event_counter']
+
+grpd = df.groupby(group_cols)
+
+rez = pd.DataFrame(columns=['user_ext', 'event_type', 'target_state', 'event_counter','ERR'])
+
+for _cols, _df in grpd:
+    srted = _df.sort_values(by='timestamp', axis=0, ascending=True).reset_index()
+    error = srted.loc[len(srted)-3,:]['ERR']
+    rez = rez.append(
+        {
+            'user_ext':_cols[0], 
+            'event_type':_cols[1], 
+            'target_state':_cols[2], 
+            'event_counter':_cols[3],
+            'error':error
+        },
+        ignore_index=True
+    )
+
+
+print(rez.groupby(['event_type','target_state'])['error'].mean())
+# event_type   target_state
+# calibration  calibrating     347.418435
+#              validating       80.452925
+# validation   validating       76.395421
+
+print(rez.groupby(['event_type','target_state'])['error'].median())
+# event_type   target_state
+# calibration  calibrating     349.082198
+#             validating       69.956001
+# validation   validating       63.249788
+
+##################################################################
+ 
+temp = df[df['target_destroyed']==True ]
+
+grpd2 = temp.groupby(group_cols)
+rez2 = pd.DataFrame(columns=['user_ext', 'event_type', 'target_state', 'event_counter'])
+for _cols, _df in grpd2:
+    srted = _df.sort_values(by='timestamp', axis=0, ascending=True).reset_index()
+    error = srted.loc[0,:]['ERR']
+    rez2 = rez2.append(
+        {
+            'user_ext':_cols[0],
+            'event_type':_cols[1],
+            'target_state':_cols[2],
+            'event_counter':_cols[3],
+            'error':error
+        },
+        ignore_index=True
+    )
+
+
+
+

### 2022-03-31 scripts/extract_validation_data.py

@@ -1,0 +1,43 @@
+import pandas as pd
+import numpy as np
+import math
+
+df = pd.read_csv("data/eye_tracking_gaze_log.csv")
+df = df[df['eye_tracking_state']=='success']
+df = df[ df['target_state']=='validating' ]
+
+# Group by the following so we get a single stream of events for each validation target
+group_cols = ['user_ext', 'event_type', 'target_state', 'event_counter']
+
+grpd = df.groupby(group_cols)
+
+rez = pd.DataFrame(columns=['target_x', 'target_y', 'gaze_x','gaze_y'])
+
+for _cols, _df in grpd:
+    srted = _df.sort_values(by='timestamp', axis=0, ascending=True).reset_index()
+    index_of_destroyed = -1
+    found = False
+    pos = -1
+    while not found:
+        if srted.loc[len(srted)+pos,:]['target_destroyed']==True:
+            index_of_destroyed = pos
+            found = True
+        pos = pos - 1
+        if len(srted)+pos < 0:
+            break
+    if found:
+        for i in range(0,5):
+            temp = srted.loc[len(srted)+pos-i,:]
+            rez = rez.append(
+               {
+                'target_x': temp['target_x'], 
+                'target_y': temp['target_y'],
+                'gaze_x': temp['gaze_x'],
+                'gaze_y': temp['gaze_y']
+               },
+               ignore_index=True
+            )
+
+rez.to_csv("data/validation_data.csv", header=True, index=False)
+ 
+

### 2022-03-31 setup.py

@@ -14,7 +14,7 @@ setup(
     name = "gazerr",
     packages = ["gazerr"],
     license = "MIT",
-    install_requires = ['numpy', 'pyyaml', 'pandas', 'fpdf'],
+    install_requires = ['numpy', 'pandas'],
     entry_points = {
         "console_scripts": ['gazerr = gazerr.cli:main']
     },

### 2022-03-31 gazerr/cli.py
Working on implementation

@@ -3,20 +3,16 @@ import argparse
 import sys
 import os
 
+from .estimate import calculate_interval
+
 ####################################################################################
 def load_calibration_data(input_path):
     df = pd.read_csv(input_path, low_memory=False)
     return df
 
 ####################################################################################
-def run_simulation(df):
-    results = pd.DataFrame()
-    locations = ["Banner Top", "Banner Bottom", "Banner Centre"]
-    for loc in locations:
-       record = {}
-       record['Location'] = loc
-       record['Error'] = 20
-       results.append(record, ignore_index=True)
+def run_simulation(df, measure, session, position):
+    results = calculate_interval(df, measure, session, position)
     return results
 
 ####################################################################################
@@ -27,22 +23,22 @@ def main():
     parser.add_argument('calibration_data',
                        metavar='calibration_data',
                        type=str,
-                       help='Path to the model calibration data contains targets and gaze coords')
+                       help='Path to the model calibration data contains targets and gaze coords - Requires columns [target_x, target_y, gaze_x, gaze_y]')
 
     parser.add_argument('measurement',
                        metavar='measurement',
                        type=float,
-                       help='Measured gaze duration.')
+                       help='Measured gaze duration [for which we want error bounds].')
 
     parser.add_argument('session_length',
                        metavar='session_length',
                        type=float,
-                       help='Measured gaze duration.')
+                       help='Length of total viewing session [Max in principle gaze duration]')
 
     parser.add_argument('ad_position',
                        metavar='ad_position',
                        type=str,
-                       help='Position of the ad in the viewport: top | mid | bottom')
+                       help='Position of the ad in the viewport. Values: top | mid | bot')
     parser.add_argument('output',
                        metavar='output',
                        type=str,
@@ -67,21 +63,8 @@ def main():
         print("    Use -f [--force] to force an overwrite over the file")
 
     df = load_calibration_data(data)
-    result = run_simulation(df)
+    result = run_simulation(df, measure, session, position)
     result.to_csv(output, index=False, header=True)
-
-#############################################################
-def print_usage(args):
-    """ Command line application usage instructions. """
-    print("USAGE ")
-    print(args[0], " <VALIDATION DATASET> <MEASUREMENT> <SESSION LENGTH> <AD POSITION> <OUTPUT>")
-    print(" <VALIDATION DATASET> - Supported file types: csv")
-    print(" <MEASUREMENT> - Measured gaze duration (seconds)")
[…]

### 2022-03-31 gazerr/estimate.py

@@ -1,0 +1,20 @@
+import pandas as pd
+import numpy as np
+import math
+
+def calculate_interval(df, measure, session, position):
+    """
+    Given a dataframe of eye tracking calibration data
+    and a gaze duration measure, session length and then
+    an ad position, then we calculate error bounds on the
+    basis of simulation.
+    """
+    if measure>session:
+        raise Exception('InvalidRequest', 'Measured duration cannot be longer than session')
+    N = 2000
+    D = math.floor(session * 10)
+    P = P = np.ndarray([D+1,D+1])
+
+    result = pd.DataFrame({"Level":[95,90,80], "Lower":[100,140,180], "Upper":[340,390,450]})
+    return result
+

### 2022-04-01 paper/paper.tex
Adding references and updating intro section of paper

@@
 and the error profile of the machine learning fixation coordinate model.
-Heretofore, gaze duration models have assumed the former situation, if indeed the possibility of error in gaze duration is even considered.
 
-In this work we develop a methodology for estimating session specific error profiles that relies on an empirical error profile of the underlying fixation model.
-The method allows us to profile whether an individual session has a compounding or cancelling error profile, and put empirical confidence bounds on the gaze duration estimates.
+The errors of eye tracking data technology can be decomposed into a range of
+independent sources that researchers need to consider \cite{Holmqvist2012}.
+Unless they are properly address these sources of error can manifest themselves
+as systematic biases across undesirable dimensions such as the age\cite{Dalrymple2018}
+or ethnicity of subjects\cite{Blignaut2013}.
+
+The use of eye fixation models to estimate gaze duration (or dwell time) in Area of
+Interest studies (AOI) was discussed by Holmqvist et al. \cite{Holmqvist2012}. The
+authors simulated the impact of gaze fixation error by adding noise accoridng to
+manufacturer specifications to data sets of low margin areas of interest. We extend 
+this idea to calculate probabilistic bounds on the error in gaze duration using
+calibration data as a source of noise distribution that is specific to both the
+study participants and the device/environment of the study. The result is an open
+source application that may be used by a wide variety of reserachers to provide
+error bounds on any gaze duration measurements.
 

### 2022-05-04 paper/paper.tex
Paper update

@@
 Eye tracking applications produce a series of gaze fixation points that can be attributed to
-objects within a subjects field of vision. Error is typically measured on the basis of individual
-gaze fixation measurements. However, often these applications are used to infer a gaze duration
-metric from a series of fixation measurements. In this work we deliver an open source application
-that estimates the error distribution of gaze duration estimates from an eye tracking validation
-log file.
+objects within a subject's field of vision. Error is typically measured on the basis of individual
+gaze fixation point measurements. However, often these applications are used to infer a gaze duration
+metric from a series of fixation measurements. In this work we develop an algorithm for estimating
+the error bounds on gaze duration measurement through monte carlo simualtion based on the content
+of an eye tracking validation log file. We provide this algorithm as an open source application
+that allows for robust estimates of the error distribution of gaze duration measurements. 
+We use this application to conduct experiments on the expected error bounds for different duration
+measurements across a range of session lengths and area of interest positions within the viewport
+of mobile devices. The results indicate...
 
@@
 
-Eye tracking applications are an important technology that permit the development of a wide 
-variety of new forms of software interaction, as well as a tool for empirical investigations 
-into human thought and behaviour. Eye tracking studies are widely used as a method of 
-measuring overt attention to visual stimuli and have been applied to study the extent to 
-which people look at, and remember, advertising \cite{Hervet2011}.
-The technology has also allowed the study of many factors that contribute to effective advertising,
-including the impact of images of faces \cite{Djamasbi2010},
-the use of animation \cite{Hamborg2012} and the relationship with
-social media posts \cite{Barreto2013}.
+Eye tracking is an important technology with a wide variety of applications. It permits evaluation
+of software user interfaces\cite{Harezlak2015}, the development of new forms of software interaction, as well as a potential method of biometric identification\cite{Kasprowski2018}. 
+Eye tracking has become a core tool for empirical investigations 
+into human behaviour and is widely used as a method of measuring explicit attention to visual stimuli. 
+It has been applied to study psychological phenomena ranging from cognition\cite{Brunye2019}
+to mental health\cite{Duque2014,Rudich-Strassler2022}, and is now routinely used to evaluate 
+advertising \cite{Hervet2011}. Eye tracking technology has allowed the marketing researchers 
+to study many factors that contribute to effective advertising, 
+including brand recall \cite{Wedel2000},
+the capture and tranfer of attention \cite{Pieters2004}, 
+the impact of images of faces \cite{Djamasbi2010},
+the attention effects of animation \cite{Hamborg2012},
+and the relationship with social media posts \cite{Barreto2013}.
 
@@
 in the training data. In addition, it has been observed that the error profile varies 
-depending on the location of the true fixation point.
+depending on the location of the true fixation point within the subjects field of view.
+The errors in fixation point measurement for a given eye tracking solution are routinely
+adjusted through a process called calibration. Improvements in the callibration process
+are an ongoing focus for the development of algorithms\cite{Zhang2014,Hassoumi2019} 
+and software tools\cite{ETCAL2018}
 
@@
 3.	The duration of the entire session of measurement.
+
 The Monte Carlo process will then repeatedly simulate sessions of the given length and gaze duration.
@@
 
+\subsection{Experiments}
+
+We evaluate the methodology by conducting experiments over a range of AOI positions and session
+lengths. The goal of which is to understand changes in the expected error profile of 
+visual attention measurement when the object position varies and under variations in the
+length of time that a subject is tracked.
+
+In both instances we use an identical validation dataset generated from a panel of 15 users
+engaged in a validation task after calibration with the SeeSo eye tracking software on mobile
+devices. In these experiments we have used all validation data as single source of error 
+estimation, but in principle these experiments could be conducted on a per user basis.
+
 \section{Results}
 
+We show all experimental results in Figure \ref{fig:plots}. The top series corresponds
+to experiments where the AOI is the top third of the viewport, the middle and bottom 
+series correspond to their respective thirds of the viewport. In each plot we show
+data for a specific session length ranging from 10 to 30 seconds, where the x axis
+depicts the measured gaze duration and the y-axis shows the confidence range for the
+true gaze duration.
+
 \section{Conclusion}
+
+Not sure yet.
 

### 2022-06-20 paper/paper.tex
Adding Bayesian posterior estimation

@@
 to mental health\cite{Duque2014,Rudich-Strassler2022}, and is now routinely used to evaluate 
-advertising \cite{Hervet2011}. Eye tracking technology has allowed the marketing researchers 
+advertising \cite{Hervet2011}. Eye tracking technology has allowed marketing researchers 
 to study many factors that contribute to effective advertising, 
@@
 When fixation models are used sequentially for the purpose of estimating gaze duration
-(as is done in advertising media studies), then there is the potential for the error
+(as is commonly done in point of salience or advertising media studies), 
+then there is the potential for the error
 to either cancel out (reducing error in gaze duration estimation) or to compound
 (increasing error in gaze duration). Which of these outcomes occurs will depend on the
-specific of the model and the conditions under which it is being used.
-Influencing factors include the ratio of true positives to true negatives in the data,
+specifics of the model and the conditions under which it is being used.
+Influencing factors include the ratio of measured gaze duration to the total viewing time
 and the error profile of the machine learning fixation coordinate model.
@@
 independent sources that researchers need to consider \cite{Holmqvist2012}.
-Unless they are properly address these sources of error can manifest themselves
+Unless they are properly addressed these sources of error can manifest themselves
 as systematic biases across undesirable dimensions such as the age\cite{Dalrymple2018}
@@
 Interest studies (AOI) was discussed by Holmqvist et al. \cite{Holmqvist2012}. The
-authors simulated the impact of gaze fixation error by adding noise accoridng to
+authors simulated the impact of gaze fixation error by adding noise according to
 manufacturer specifications to data sets of low margin areas of interest. We extend 
@@
 on the basis of the gaze fixation errors on the validation data from eye tracking
-fixation model.
-This error profile can be stratified into a grid across the screen dimensions.
-For each of these grid locations we calculate two error probabilities, the false negative probability (when true fixation is on the grid and the model produces coordinates outside the grid.
+fixation model. This error profile can be stratified into a grid across the screen dimensions. 
+For each of these grid locations we calculate two error probabilities, the false negative 
+probability (when true fixation is on the grid and the model produces coordinates outside the grid. 
 And the false positive probability, when the true fixation is outside the grid location and the model produces a prediction inside the grid square.
@@
 
-In the first stage of the methodology we provide an estimation of the distribution of 
-measurements for a given true gaze duration. These distributions are produced at multiple locations
-across the screen, so that we have a set of distributions, that captures the expected
-variation in measurement depending on the length of and locus of true fixation.
+In the first stage of the methodology we provide an estimate of the distribution of measured
+durations for all possible true gaze durations. These distributions are produced by generating
+random fixation paths across the available screen dimensions, and adding noise that is 
+drawn from the calibration data to make it consistent with the observed properties of the gaze model.
+This produces a set of distributions that capture the expected
+variation in measurement depending on the length of true fixation.
 
-We want to estimate a the probability distribution over measured durations $\hat{d}$ given a
-known duration $d$, the session length $s$ and the ad location $l$. 
+Formally stated this is the probability distribution over measured durations $\hat{d}$ given a
+known duration $d$, the session length $s$ and the gaze target location $l$. 
 We express this distribution as $P(\hat{d}|d,l,s)$. The
@@
 These points determine the $\delta_x$ and $\delta_y$ error in the measurement
-of gaze fixation point $f$. We draw these coordinates from the validation file such the
+of gaze fixation point $f$. We draw these coordinates from the Calibration file such the
 the probability of the point being chosen is propotional to the distance of its target
@@
 \begin{algorithmic}
-  \Require $\hat{d},l,s,C$
+  \Require $l,s,C$
   \Ensure $P(\hat{d}|d,l,s)$
-  \State $N \gets 2000$
-  \State $D \gets floor(s \times 10)$ \Comment{Durations sampled at 100ms intervals}
-  \State $P \gets Array(D+1,D+1)$ \Comment{Result is a 2 dimensional array}
-  \If{$\hat{d} > s$}
-    \State error
-  \EndIf
-  \For{$d \in 0 to D$} \Comment{Iterate over all D+1 possible durations}
-    \For{$n \in 1 to N$} \Comment{Collect N samples}
+  \State $N \gets samples$              \Comment{Configure simulation samples per duration}
+  \State $D \gets floor(s / increment)$ \Comment{Durations are sampled at discrete intervals}
+  \State $P \gets Array(D+1,D+1)$       \Comment{The distribution P is a 2 dimensional array}
+  \For{$d \in 0 to D$}                  \Comment{Iterate over all possible true gaze durations}
+    \For{$n \in 0 to N$}                \Comment{Iterature to collect the N samples for d}
       \State $p \gets generatePath(d,l,s)$
       \State $d_e \gets generateMeasurement(p,C)$  
-      \State $P[d,d_e] += 1/2000$  
+      \State $P[d,d_e] += 1/N$          \Comment{Probability increment for a measuremnt of $d_e$}
     \EndFor
   \EndFor
+  \State return P
 \end{algorithmic}
@@
 The bulk of the work is done by two functions inside the simulation loop. The first function
-$generatePath(d,l,s)$ takes a true gaze duration, position of ad and viewing session length and
-generates a path of fixation points compatible with those parameters. The second function 
-$generateMeasurement(p,C)$, takes the path and the eye tracking calibration data and produces
-an sample of measured attention time. Repeated application of this process over different 
+$generatePath(d,l,s)$ takes a true gaze duration (for that simulation iteration) the location
+of the gaze target and the viewing session length. It then
+generates a random path of fixation points that is compatible with those parameters.
+The second function 
+$generateMeasurement(p,l,C)$, takes the path, target location and the eye tracking calibration 
+data and produces a sample of measured attention time. It does this by using the calibration data to sample
+fixation errors and apply them to the path data. The noisy path data is then used to determine
+the estimated gaze duration by looking at the number of fixation points that intersect with the target
+location.
+Repeated application of this process for a given value of $d$ produces multiple samples of $\hat{d}$
+for that true gaze duration. Repeating the process over different 
 values of $d$ allows us to populate a two dimensional array where the first dimension is the
-true duration and the second the estimated. Note: that our probability distribution is discrete
-because we restrict ourselves to sampling every 100ms point between zero and the session length.
+true duration and the second the estimated duration. Note that our probability distribution is discrete
+because we restrict ourselves to sampling every increment point between zero and the session length.
 

### 2022-06-20 gazerr/cli.py

@@ -49,12 +49,12 @@ def main():
     parser.add_argument('measurement',
                        metavar='measurement',
                        type=float,
-                       help='Measured gaze duration [for which we want error bounds].')
+                       help='Measured gaze duration (in milliseconds) [for which we want error bounds].')
 
     parser.add_argument('session_length',
                        metavar='session_length',
                        type=float,
-                       help='Length of total viewing session [Max in principle gaze duration]')
+                       help='Length of total viewing session (in milliseconds) [Max in principle gaze duration]')
 
     parser.add_argument('target_top_left',
                        metavar='target_top_left',

### 2022-06-20 gazerr/estimate.py

@@ -8,8 +8,8 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
     Calculate error bounds on a gaze duration measurement.
     Params
     * df: A dataframe of eye tracking calibration data
-    * measure: A gaze duration measurement
-    * session: The eye tracking session length
+    * measure: A gaze duration measurement (in milliseconds)
+    * session: The eye tracking session length (in milliseconds)
     * top_l_x, top_l_y, bot_r_x, bot_r_y : Top left and bottom right coordinates of target.
     
     Returns: A series of probabilistic error bounds
@@ -25,11 +25,12 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
     # Force measure and session to be rounded integers
     measure = round(float(measure))
     session = round(float(session))
-    increment = 25
-    N = 600
+    increment = 100
+    N = 1000
     inc = 1/N
     D = math.floor(session / increment) + 1
     G = math.floor(measure / increment) + 1
+    prior = 1/D
     P = np.zeros([D,D])
     top_l = (top_l_x, top_l_y)
     bot_r = (bot_r_x, bot_r_y)
@@ -79,15 +80,45 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
             dhat = sum(insiders)
             P[d,dhat] += inc    
 
-    # At this point in the algorithm we have a set of probability distributions over dhat
-    # for a range of potential values of true gaze duration: d
+    # #######################################################################################
+    # At this point in the algorithm we have a set of probability distributions over
+    # the measured duration (dhat) for a range of potential values of true gaze duration (d)
+    # This is the distribution p(dhat|d)
+    #
+    # We need to invert this to get p(d|dhat) using Bayes rule. 
 
-    # As an intermediate step for testing, lets return distrubution around measurement = d
-    result = extract_intervals(P[G,:], [0.99, 0.95,0.90,0.80], increment=increment )
+    posterior = invert_distribution(prior, P)
 
-    #print(P[G,:])
-    #print("Total Probability:", P[G,:].sum())
+    # Return distrubution around measurement = d
+    result = extract_intervals(posterior[G,:], [0.99, 0.95,0.90,0.80], increment=increment )
+
+    print(posterior[G,:])
+    print("Total Probability:", posterior[G,:].sum())
+
     return result
+
+###########################################################
+def invert_distribution(prior, likelihood):
+    """
[…]

### 2022-06-27 paper/paper.tex
Code prior to refactor

@@
 objects within a subject's field of vision. Error is typically measured on the basis of individual
-gaze fixation point measurements. However, often these applications are used to infer a gaze duration
-metric from a series of fixation measurements. In this work we develop an algorithm for estimating
-the error bounds on gaze duration measurement through monte carlo simualtion based on the content
+gaze fixation point measurements. These applications are often used to infer a gaze duration
+metric from a series of fixation measurements. There is no direct method for infering the error in
+a gaze duration measurement from an error in fixation points. 
+In this work we develop an algorithm for estimating
+the error bounds on gaze duration measurement through monte carlo simulation based on the content
 of an eye tracking validation log file. We provide this algorithm as an open source application
@@
 measurements across a range of session lengths and area of interest positions within the viewport
-of mobile devices. The results indicate...
+of mobile devices. The results indicate that error in gaze duration estimation is sensitive to
+the size of fixation error as well as the proportion of total field of view that the object occupies
+and the proportion of viewing time that was taken with gaze fixation.
 

### 2022-06-27 gazerr/cli.py

@@ -11,7 +11,7 @@ def load_calibration_data(input_path):
     return df
 
 ####################################################################################
-def run_simulation(df, measure, session, top_left, bot_right):
+def run_simulation(df, measure, session, top_left, bot_right, posterior):
     top = top_left.split(",")
     bot = bot_right.split(",")
     if len(top) != 2:
@@ -66,12 +66,18 @@ def main():
                        type=str,
                        help='X,Y Position for bottom right of target bounding box.')
 
+    parser.add_argument('posterior_file',
+                       metavar='posterior_file',
+                       type=str,
+                       help='Path to write out the posterior distribution.')
+
     args = parser.parse_args()
     data = args.calibration_data
     measure = args.measurement
     session = args.session_length
     top_left = args.target_top_left
     bottom_right = args.target_bottom_right
+    posterior = args.posterior_file
 
     if not os.path.isfile(data):
         print(" ERROR")
@@ -79,7 +85,7 @@ def main():
         sys.exit()
 
     df = load_calibration_data(data)
-    result = run_simulation(df, measure, session, top_left, bottom_right)
+    result = run_simulation(df, measure, session, top_left, bottom_right, posterior)
     print(result)
 
 ##########################################################################################

### 2022-06-27 gazerr/estimate.py

@@ -25,11 +25,11 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
     # Force measure and session to be rounded integers
     measure = round(float(measure))
     session = round(float(session))
-    increment = 100
-    N = 1000
+    increment = 50
+    N = 500
     inc = 1/N
     D = math.floor(session / increment) + 1
-    G = math.floor(measure / increment) + 1
+    G = math.floor(measure / increment) 
     prior = 1/D
     P = np.zeros([D,D])
     top_l = (top_l_x, top_l_y)
@@ -73,7 +73,7 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
     for d in range(0,D):
         for n in range(0,N):
             in_path = get_path(d, top_l_x, top_l_y, bot_r_x, bot_r_y)
-            out_path = get_path(D-d-1, 0, 0, max_X, max_y, top_l_x, top_l_y, bot_r_x, bot_r_y)
+            out_path = get_path(D-d, 0, 0, max_X, max_y, top_l_x, top_l_y, bot_r_x, bot_r_y)
             path = in_path + out_path
             measured_path = [apply_measurement_noise(point) for point in path] 
             insiders = [int(inside(p,top_l,bot_r)) for p in measured_path]
@@ -89,11 +89,14 @@ def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y)
 
     posterior = invert_distribution(prior, P)
 
-    # Return distrubution around measurement = d
+    # Return distrubution for Measurement G
     result = extract_intervals(posterior[G,:], [0.99, 0.95,0.90,0.80], increment=increment )
 
-    print(posterior[G,:])
-    print("Total Probability:", posterior[G,:].sum())
+    measurements = [x * increment for x in range(0,D)]
+    print("Posterior\n", posterior[G,:])
+    print("Total Probability: ", posterior[G,:].sum())
+    expected_value = (posterior[G,:]*measurements).sum()
+    print("Expected Value: ", expected_value) 
 
     return result
 
@@ -114,26 +117,36 @@ def invert_distribution(prior, likelihood):
     for dhat in range(0,dim_measured):
         # Extract those measurements as a vector
         vec = likelihood[:,dhat]
-        numera = prior * vec
+        numera = np.nan_to_num(prior * vec)
         denom = numera.sum()
[…]

### 2022-06-27 scripts/generate_datasets.py

@@ -1,0 +1,49 @@
+import math
+import random
+import pandas as pd
+
+header = ["target_x","target_y","gaze_x","gaze_y"]
+
+maes = [25, 50, 75, 100, 125, 150]
+
+# We use an iPhone 7 : which has a viewport of 375px × 667
+#  the number of software pixels (CSS pixels) present on a screen. 
+device_width = 375
+device_height = 667
+
+N = 1000
+
+def add_noise(x,y,maerr):
+    factor = 1.85
+    part = math.sqrt(math.pow(maerr,2)/2)
+    x_err = random.uniform(-part*factor,part*factor)
+    y_err = random.uniform(-part*factor,part*factor)
+    return x+x_err,y+y_err
+
+def euclidean(row):
+    return math.sqrt(
+       math.pow(row["target_x"] - row["gaze_x"], 2) +
+       math.pow(row["target_y"] - row["gaze_y"], 2)
+    )
+
+for mae in maes:
+    results = pd.DataFrame(columns=header)
+    filename = "data/validation_%s_MAE.csv"%str(mae)
+    for n in range(N):
+        my_x = int(random.uniform(0,device_width))
+        my_y = int(random.uniform(0,device_height))
+        gaze_x, gaze_y = add_noise(my_x,my_y,mae)
+        results = results.append({
+              "target_x":my_x,
+              "target_y":my_y,
+              "gaze_x":gaze_x,
+              "gaze_y":gaze_y
+           }, ignore_index=True
+        )
+    results.to_csv(filename, index=False,header=True)
+    results['error'] = results.apply(euclidean, axis=1)
+    print("Expected MAE: ", mae)
+    print("Simulated MAE: ", results['error'].mean())
+
+
+

### 2022-06-27 tests/test_functions.py

@@ -1,0 +1,59 @@
+import sys
+import pytest
+import pandas as pd
+import numpy as np
+
+"""
+   Signature of functions to test
+   def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y):
+   def invert_distribution(prior, likelihood):
+   def extract_intervals(dist, intervals, increment):
+   def extract_screen_limits(df, bot_r_x, bot_r_y):
+   def get_path(n, l_x, l_y, r_x, r_y, exc_l_x=-1,exc_l_y=-1,exc_r_x=-1,exc_r_y=-1):
+   def point_inside(temp_x, temp_y, exc_l_x, exc_l_y, exc_r_x, exc_r_y):
+   def inside(point, top_l, bot_r):
+"""
+from gazerr.estimate import invert_distribution
+from gazerr.estimate import get_path
+from gazerr.estimate import point_inside
+from gazerr.estimate import inside
+from gazerr.estimate import extract_intervals
+
+def test_pathno_excl():
+    mypath = get_path(10, 0, 0, 100, 100)
+    assert len(mypath) == 10
+    for p in mypath:
+        assert p[0] >= 0.0
+        assert p[1] >= 0.0
+        assert p[0] <= 100.0
+        assert p[1] <= 100.0
+
+def test_inside():
+    inone = inside((30,30), (20,20), (40,40) )
+    outone = inside((0,0), (20,20), (40,40) )
+    assert inone == True
+    assert outone == False
+
+def test_point_inside():
+    inone = point_inside(30,30, 20,20, 40,40 )
+    outone = point_inside(0,0, 20,20, 40,40 )
+    assert inone == True
+    assert outone == False
+
+def test_calculate_interval():
+    prior = 0.25
+    likelihood = np.array([[0.1,0.1,0.3,0.5],[0.6,0.2,0.1,0.1],[0.2,0.5,0.2,0.1],[0.1,0.1,0.7,0.1]])
+    posterior = invert_distribution(prior, likelihood)
+    result = extract_intervals(posterior[1,:], [0.95,0.90], increment=100 )
+    for r in result.iterrows():
+       assert r[1]['Lower'] <= r[1]['Upper']
+
+
+def test_invert_distribution():
+    prior = 0.25
+    likelihood = np.array([[0.1,0.1,0.3,0.5],[0.6,0.2,0.1,0.1],[0.2,0.5,0.2,0.1],[0.1,0.1,0.7,0.1]])
[…]

### 2022-06-27 gazerr/cli.py
CLI and estimation refactor - generate and write expectations

@@ -4,6 +4,8 @@ import sys
 import os
 
 from .estimate import calculate_interval
+from .estimate import calculate_posterior
+from .estimate import calculate_expectations
 
 ####################################################################################
 def load_calibration_data(input_path):
@@ -11,7 +13,7 @@ def load_calibration_data(input_path):
     return df
 
 ####################################################################################
-def run_simulation(df, measure, session, top_left, bot_right, posterior):
+def run_simulation(df, measure, session, top_left, bot_right, results_path):
     top = top_left.split(",")
     bot = bot_right.split(",")
     if len(top) != 2:
@@ -33,8 +35,13 @@ def run_simulation(df, measure, session, top_left, bot_right, posterior):
         print("BOTTOM RIGHT COORDINATES MUST BE A COMMA SEPARATED PAIR OF INTEGERS")
         exit(1)
 
-    results = calculate_interval(df, measure, session, tlx, tly, brx, bry)
-    return results
+    increment = 50
+    posterior = calculate_posterior(df, session, increment, tlx, tly, brx, bry)
+    expected = calculate_expectations(session, increment, posterior)
+    expected.to_csv(results_path+"/expected_values.csv",index=False, header=True)
+    results = calculate_interval(measure, session, increment, posterior)
+    results.to_csv(results_path+"/error_bounds.csv",index=False, header=True)
+
 
 ####################################################################################
 def main():
@@ -66,10 +73,10 @@ def main():
                        type=str,
                        help='X,Y Position for bottom right of target bounding box.')
 
-    parser.add_argument('posterior_file',
-                       metavar='posterior_file',
+    parser.add_argument('results',
+                       metavar='results',
                        type=str,
-                       help='Path to write out the posterior distribution.')
+                       help='Path to directory for results: posterior distribution and expected values.')
 
     args = parser.parse_args()
     data = args.calibration_data
@@ -77,7 +84,12 @@ def main():
     session = args.session_length
     top_left = args.target_top_left
     bottom_right = args.target_bottom_right
-    posterior = args.posterior_file
+    results = args.results
+
+    if measure > session:
+        print(" ERROR")
+        print('Measured duration cannot be longer than session')
+        sys.exit()
 
     if not os.path.isfile(data):
         print(" ERROR")
@@ -85,8 +97,9 @@ def main():
         sys.exit()
 
     df = load_calibration_data(data)
-    result = run_simulation(df, measure, session, top_left, bottom_right, posterior)
-    print(result)
+
[…]

### 2022-06-27 gazerr/estimate.py

@@ -2,40 +2,75 @@ import pandas as pd
 import numpy as np
 import random
 import math
+ 
+#################################################################################
+def calculate_interval( measure, session, increment, posterior):
+    measure = round(float(measure))
+    session = round(float(session))
+    D = math.floor(session / increment) + 1
+    G = math.floor(measure / increment) 
+    if measure > session:
+        raise Exception('InvalidRequest', 'Measured duration cannot be longer than session')
+    # Return distrubution for Measurement G
+    result = extract_intervals(posterior[G,:], [0.99, 0.95,0.90,0.80], increment=increment )
+    measurements = [x * increment for x in range(0,D)]
+    print("Posterior\n", posterior[G,:])
+    print("Total Probability: ", posterior[G,:].sum())
+    expected_value = (posterior[G,:]*measurements).sum()
+    print("Expected Value: ", expected_value)
+    return result
 
-def calculate_interval(df, measure, session, top_l_x, top_l_y, bot_r_x, bot_r_y):
+#################################################################################
+def calculate_expectations(session, increment, posterior):
+    session = round(float(session))
+    D = math.floor(session / increment) + 1
+    measurements = [x * increment for x in range(0,D)]
+    results = pd.DataFrame(columns=["Measured","Expected"])
+    for d in range(0,D):
+        expected_value = (posterior[d,:]*measurements).sum()
+        print("Measured:", measurements[d], " Expected Value: ", expected_value)
+        results = results.append({
+            "Measured":measurements[d],
+            "Expected":expected_value
+        }, ignore_index=True)
+
+    return results
+
+#################################################################################
+def calculate_posterior(df, session, increment, top_l_x, top_l_y, bot_r_x, bot_r_y, max_x=None, max_y=None):
     """
-    Calculate error bounds on a gaze duration measurement.
+    Calculate the posterior distributions for actual gaze duration over all possible measurements
+    For a given session length, and granuality specified by interval.
     Params
     * df: A dataframe of eye tracking calibration data
-    * measure: A gaze duration measurement (in milliseconds)
     * session: The eye tracking session length (in milliseconds)
+    * interval: The discrete step size for possible measurements (in milliseconds)
[…]

### 2022-06-27 scripts/RUN_EXPERIMENTS.sh

@@ -1,0 +1,20 @@
+
+mkdir results/MREC_MAE_25
+python ./gazerr-runner.py data/validation_25_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_25
+
+mkdir results/MREC_MAE_50
+python ./gazerr-runner.py data/validation_50_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_50
+
+mkdir results/MREC_MAE_75
+python ./gazerr-runner.py data/validation_75_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_75
+
+mkdir results/MREC_MAE_100
+python ./gazerr-runner.py data/validation_100_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_100
+
+mkdir results/MREC_MAE_125
+python ./gazerr-runner.py data/validation_125_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_125
+
+mkdir results/MREC_MAE_150
+python ./gazerr-runner.py data/validation_150_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_150
+
+

### 2022-06-27 paper/paper.tex
Paper update to include initial plots

@@
 
-We evaluate the methodology by conducting experiments over a range of AOI positions and session
-lengths. The goal of which is to understand changes in the expected error profile of 
-visual attention measurement when the object position varies and under variations in the
-length of time that a subject is tracked.
+We apply the gazerr application to investigate the relationship between measured gaze duration
+and expected real duration under a variety of changing scenarios. These scenarios are variations
+in the underlying error indicated by the eye tracking validation file and length of the session
+in which the user could be looking at the area of interest.
 
-In both instances we use an identical validation dataset generated from a panel of 15 users
-engaged in a validation task after calibration with the SeeSo eye tracking software on mobile
-devices. In these experiments we have used all validation data as single source of error 
-estimation, but in principle these experiments could be conducted on a per user basis.
+For the sake of our simulations we use a fixed sized simulated device with viewport pixel 
+dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium rectangle ad unit (MREC)
+that is located toward the top of the screen. We iterate over a range of potential mean error
+values (determined to be the mean Euclidean distance between the measured fixation and actual fixation).
+For each of these mean error values we generate a synthetic eye tracking validation file.
+
+We feed each of these synthetic validation files into gazerr to calculate posterior probability
+distributions over true gaze duration for the set of measured gaze durations. These posterior
+distributions are converted into expected values of true gaze duration for each measured duration.
+We repeat the above process for multiple session lengths to investigate the impact of increasing
+session length on the expected gaze duration for a fixed set of measurements.
+
+All source code for these experiments is available in the gazerr open source package. 
+Raw data is generated by running the $RUN\_EXPERIMENTS.sh$ bash script. These results
+are then analysed and used to generatedt the resulting plots with the script 
+$ANALYSE\_RESULTS.sh$.
 
 \section{Results}
+
+The first experimental result is the expected true duration against the measured duration
+for a single eye tracking application with a mean fixation error of 50 pixels. These
+results are shown in Figure \ref{fig:measured_vs_expected}.
+
+\begin{figure}[h!]
+\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
+\caption{Measured versus expected gaze duration. MREC in 10s Session with 50px Mean Error.}
+\label{fig:measured_vs_expected}
+\end{figure}
+
+The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we see under
+multiple variations of the simulation configuration. At lower meaurement values there is a 
+tendency for the measurement to be an over estimation, and at higher values an under-estimation.
+
+We next examined the way that exected error in the gaze duration estimate tracks with expected 
+error in the gaze fixation points. This involved calculating the mean absolute error in gaze
+duration for each of the mean fixation error validation files. We performed this calculation
+using two methods. The first method, called uniform, assumes that all measured values are 
+equally likely. The second method, called smoothed, assumes that low measurements are much
+more likely than higher measurement, and that this drop in likelihood follows an approximately
+exponential distribution.
+
+\begin{figure}[h!]
+\includegraphics[scale=0.5]{../results/Fixation_vs_duration_error.png}
+\caption{Fixation error versus gaze duration error.}
+\label{fig:fixation_vs_duration}
+\end{figure}
+
+As shown in Figure \ref{fig:fixation_vs_duration} we see
 
@@
 
-Not sure yet.
-
 \bibliographystyle{IEEEtran}

### 2022-06-27 gazerr/cli.py

@@ -1,4 +1,5 @@
 import pandas as pd
+import numpy as np
 import argparse
 import sys
 import os
@@ -37,6 +38,7 @@ def run_simulation(df, measure, session, top_left, bot_right, results_path):
 
     increment = 50
     posterior = calculate_posterior(df, session, increment, tlx, tly, brx, bry)
+    np.savetxt(results_path+"/posterior.out", posterior)
     expected = calculate_expectations(session, increment, posterior)
     expected.to_csv(results_path+"/expected_values.csv",index=False, header=True)
     results = calculate_interval(measure, session, increment, posterior)

### 2022-06-27 gazerr/estimate.py

@@ -61,7 +61,7 @@ def calculate_posterior(df, session, increment, top_l_x, top_l_y, bot_r_x, bot_r
 
     # Force session to be rounded integers
     session = round(float(session))
-    N = 500
+    N = 1500
     inc = 1/N
     D = math.floor(session / increment) + 1
     prior = 1/D
@@ -108,20 +108,18 @@ def calculate_posterior(df, session, increment, top_l_x, top_l_y, bot_r_x, bot_r
     for d in range(0,D):
         for n in range(0,N):
             in_path = get_path(d, top_l_x, top_l_y, bot_r_x, bot_r_y)
-            out_path = get_path(D-d, 0, 0, max_x, max_y, top_l_x, top_l_y, bot_r_x, bot_r_y)
+            out_path = get_path(D-d-1, 0, 0, max_x, max_y, top_l_x, top_l_y, bot_r_x, bot_r_y)
             path = in_path + out_path
             measured_path = [apply_measurement_noise(point) for point in path] 
             insiders = [int(inside(p,top_l,bot_r)) for p in measured_path]
             dhat = sum(insiders)
             P[d,dhat] += inc    
 
-    # #######################################################################################
-    # At this point in the algorithm we have a set of probability distributions over
-    # the measured duration (dhat) for a range of potential values of true gaze duration (d)
-    # This is the distribution p(dhat|d)
-    #
-    # We need to invert this to get p(d|dhat) using Bayes rule. 
-
+    # ########################################################################
+    # At this point we have a set of probability distributions over
+    # the measured duration (dhat) for a range of potential values 
+    # of true gaze duration (d). This is the distribution p(dhat|d)
+    # We invert this to get p(d|dhat) using Bayes rule. 
     posterior = invert_distribution(prior, P)
 
     return posterior

### 2022-06-27 scripts/ANALYSE_RESULTS.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+# NOTE: This script depends on the output of RUN_EXPERIMENTS.sh
+
+# Generate the Measured Versus Expected Plot
+python plot_measured_vs_expected.py
+
+# Generate the fixation versus duration error
+python scripts/plot_expected_errors.py
+

### 2022-06-27 scripts/RUN_EXPERIMENTS.sh

@@ -1,3 +1,4 @@
+#!/bin/bash
 
 mkdir results/MREC_MAE_25
 python ./gazerr-runner.py data/validation_25_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_25

### 2022-06-27 scripts/plot_expected_errors.py

@@ -1,0 +1,44 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import expon
+import matplotlib.pyplot as plt
+
+file_prefix = "results/MREC_MAE_"
+filename = "/expected_values.csv"
+
+errors = [25,50,75,100,125,150]
+gaze_err_uni = []
+gaze_err_exp = []
+
+# This creates the weighting for measurements
+x = expon.rvs(scale=30, size=10000)
+vals = plt.hist(x, density=False, edgecolor='black', bins=21)
+vallist = vals[0].tolist() 
+dist_exp = [x / sum(vallist) for x in vallist]
+dist_uni = [1/len(vallist) for x in vallist]
+
+for err in errors:
+    file = file_prefix + str(err) + filename
+    df = pd.read_csv(file)
+    df['error'] = abs(df['Measured']-df['Expected'])
+    df['uni'] = dist_uni
+    df['exp'] = dist_exp
+    df['err_uni'] = df['error']*df['uni'] 
+    df['err_exp'] = df['error']*df['exp'] 
+    gaze_err_uni.append(df['err_uni'].sum())
+    gaze_err_exp.append(df['err_exp'].sum()) 
+
+fig = plt.figure(figsize=(10,10))
+
+ax = fig.add_subplot()
+
+ax.set_title('Fixation Vs Duration Expected Error')
+ax.plot(errors, gaze_err_uni, c='darkorange', label="Uniform")
+ax.plot(errors, gaze_err_exp, c='teal', label="Exponential")
+ax.set_xlabel('Gaze Fixation μ Error (px)', fontsize=15)
+ax.set_ylabel('Gaze Duration μ Error (ms)', fontsize=15)
+ax.set_xlim([0, 175])
+ax.legend()
+
+plt.savefig("results/Fixation_vs_duration_error.png")
+

### 2022-06-27 scripts/plot_measured_vs_expected.py

@@ -1,0 +1,21 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+file = "results/MREC_MAE_50/expected_values.csv"
+df = pd.read_csv(file)
+
+measured = df['Measured']
+expected = df['Expected']
+
+plt.figure(figsize=(10,10))
+plt.scatter(measured, expected, c='crimson')
+
+p1 = max(max(measured), max(expected))
+p2 = min(min(measured), min(expected))
+plt.plot([p1, p2], [p1, p2], 'b-')
+plt.xlabel('Measured Gaze Duration', fontsize=15)
+plt.ylabel('Expected Gaze Duration', fontsize=15)
+plt.axis('equal')
+
+plt.savefig("results/Measured_vs_expected.png")
+

### 2022-06-29 paper/paper.tex
Changing density of experiments

@@
 
-As shown in Figure \ref{fig:fixation_vs_duration} we see
+As shown in Figure \ref{fig:fixation_vs_duration} we see that the error in gaze duration
+grows with error in fixation points, as would be expected. However, when we use the exponential
+weighting to calculate the expected error in duration, the gaze duration error grows much slower
+than in the uniform case. This suggests that the observed tendency toward smaller measurements
+in area of interest studies means lower expected error.
 
-We show all experimental results in Figure \ref{fig:plots}. The top series corresponds
-to experiments where the AOI is the top third of the viewport, the middle and bottom 
-series correspond to their respective thirds of the viewport. In each plot we show
-data for a specific session length ranging from 10 to 30 seconds, where the x axis
-depicts the measured gaze duration and the y-axis shows the confidence range for the
-true gaze duration.
+In our final experiment we look at the effect of both the size of the measured 
+gaze duration and the mean error of fixation on the expected error in gaze duration.
 
 \section{Conclusion}
+
+We have demonstrated that the statistial information contained in eye tracking validation
+data can be utilised for estimating errors in gaze duration estimation. The process involves
+simulating the distribution of measurements for a given true durastion, and then inverting
+it into a distribution over true duration for a given measurement.
+
+We have released this approach as the open source application gazerr, and used it to explore
+the relationship between fixation error and duration error. Our results show that the nature
+of the error in gaze duration depends on the size of the measurement. Smaller measurements
+tend toward over prediction, and larger measurements towards under prediction. We plotted
+the relationship between expected error and saw that observed tendency toward smaller duration
+measurements (estimated with an psuedo-exponential distribution) delivers a slower growing
+expected duraiton error with the mean error in fixation points.
+
+Finally, our simualtions revealed that for large error in fixation points certain measurements
+become some unlikely that our monte carlo simualtions placed none of the probability density
+in those regions. This strongly suggests that an additional downside to large fixation point error
+is a reduced range of measurement for area of interest studies.
 

### 2022-06-29 gazerr/estimate.py

@@ -61,7 +61,7 @@ def calculate_posterior(df, session, increment, top_l_x, top_l_y, bot_r_x, bot_r
 
     # Force session to be rounded integers
     session = round(float(session))
-    N = 1500
+    N = 1000
     inc = 1/N
     D = math.floor(session / increment) + 1
     prior = 1/D

### 2022-06-29 scripts/ANALYSE_RESULTS.sh

@@ -3,7 +3,7 @@
 # NOTE: This script depends on the output of RUN_EXPERIMENTS.sh
 
 # Generate the Measured Versus Expected Plot
-python plot_measured_vs_expected.py
+python scripts/plot_measured_vs_expected.py
 
 # Generate the fixation versus duration error
 python scripts/plot_expected_errors.py

### 2022-06-29 scripts/RUN_EXPERIMENTS.sh

@@ -1,21 +1,48 @@
 #!/bin/bash
 
-mkdir results/MREC_MAE_25
-python ./gazerr-runner.py data/validation_25_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_25
+mkdir results/MREC_MAE_20
+python ./gazerr-runner.py data/validation_20_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_20
+
+mkdir results/MREC_MAE_30
+python ./gazerr-runner.py data/validation_30_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_30
+
+mkdir results/MREC_MAE_40
+python ./gazerr-runner.py data/validation_40_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_40
 
 mkdir results/MREC_MAE_50
 python ./gazerr-runner.py data/validation_50_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_50
 
-mkdir results/MREC_MAE_75
-python ./gazerr-runner.py data/validation_75_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_75
+mkdir results/MREC_MAE_60
+python ./gazerr-runner.py data/validation_60_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_60
+
+mkdir results/MREC_MAE_70
+python ./gazerr-runner.py data/validation_70_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_70
+
+mkdir results/MREC_MAE_80
+python ./gazerr-runner.py data/validation_80_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_80
+
+mkdir results/MREC_MAE_90
+python ./gazerr-runner.py data/validation_90_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_90
 
 mkdir results/MREC_MAE_100
 python ./gazerr-runner.py data/validation_100_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_100
 
-mkdir results/MREC_MAE_125
-python ./gazerr-runner.py data/validation_125_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_125
+mkdir results/MREC_MAE_110
+python ./gazerr-runner.py data/validation_110_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_110
+
+mkdir results/MREC_MAE_120
[…]

### 2022-06-29 scripts/generate_datasets.py

@@ -4,14 +4,15 @@ import pandas as pd
 
 header = ["target_x","target_y","gaze_x","gaze_y"]
 
-maes = [25, 50, 75, 100, 125, 150]
+maes = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
 
-# We use an iPhone 7 : which has a viewport of 375px × 667
-#  the number of software pixels (CSS pixels) present on a screen. 
+# We roughly approximate an iPhone 7 
+# Which has a viewport dimensions of 375px × 667
+#  (number of software pixels (CSS pixels) present on a screen) 
 device_width = 375
 device_height = 667
 
-N = 1000
+N = 5000
 
 def add_noise(x,y,maerr):
     factor = 1.85
@@ -30,8 +31,8 @@ for mae in maes:
     results = pd.DataFrame(columns=header)
     filename = "data/validation_%s_MAE.csv"%str(mae)
     for n in range(N):
-        my_x = int(random.uniform(0,device_width))
-        my_y = int(random.uniform(0,device_height))
+        my_x = random.randrange(0, device_width, 5)
+        my_y = random.randrange(0, device_height, 5)
         gaze_x, gaze_y = add_noise(my_x,my_y,mae)
         results = results.append({
               "target_x":my_x,

### 2022-06-29 scripts/plot_expected_errors.py

@@ -6,7 +6,7 @@ import matplotlib.pyplot as plt
 file_prefix = "results/MREC_MAE_"
 filename = "/expected_values.csv"
 
-errors = [25,50,75,100,125,150]
+errors = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
 gaze_err_uni = []
 gaze_err_exp = []
 

### 2022-06-29 gazerr/estimate.py
Switching data generation to a grid of points, making noise sampling drawn from top3 closest points in the monte carlo process

@@ -95,11 +95,16 @@ def calculate_posterior(df, session, increment, top_l_x, top_l_y, bot_r_x, bot_r
        temp.sort_values('distance', inplace=True)
        top1 = temp.loc[0,:]['distance']
        top2 = temp.loc[1,:]['distance']
-       threshold = top2/(top1+top2)
-       if random.uniform(0,1) < threshold:
+       top3 = temp.loc[2,:]['distance']
+       threshold1 = (top2+top3)/(top1+top2+top3)
+       threshold2 = threshold1 + (1-threshold1)*(top3)/(top2+top3)
+       randy = random.uniform(0,1)
+       if randy < threshold1:
            noise_set = temp.loc[0,:]
-       else:
+       elif randy < threshold2:
            noise_set = temp.loc[1,:]
+       else:
+           noise_set = temp.loc[2,:]
        recs = len(noise_set['err_x'])
        rn = random.randrange(0,recs)
        return (point[0] + noise_set['err_x'][rn], point[1] + noise_set['err_y'][rn])

### 2022-06-29 scripts/generate_datasets.py

@@ -12,7 +12,7 @@ maes = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
 device_width = 375
 device_height = 667
 
-N = 5000
+N = 1500
 
 def add_noise(x,y,maerr):
     factor = 1.85
@@ -31,8 +31,8 @@ for mae in maes:
     results = pd.DataFrame(columns=header)
     filename = "data/validation_%s_MAE.csv"%str(mae)
     for n in range(N):
-        my_x = random.randrange(0, device_width, 5)
-        my_y = random.randrange(0, device_height, 5)
+        my_x = random.randrange(0, device_width, 20)
+        my_y = random.randrange(0, device_height, 20)
         gaze_x, gaze_y = add_noise(my_x,my_y,mae)
         results = results.append({
               "target_x":my_x,

### 2022-06-29 scripts/plot_error_surface.py
Adding anaysis scripts and updates

@@ -1,0 +1,45 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import expon
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import matplotlib as mpl
+
+file_prefix = "results/MREC_MAE_"
+filename = "/expected_values.csv"
+
+errors = [25,50,75,100,125,150]
+#errors = [20,40,60,80,100,120,140,160]
+
+D = 21
+dim1 = []
+dim2 = []
+dim3 = []
+
+def generate_2d_surface_plot(dim1, dim2, dim3, colone, coltwo, value):
+    mpl.rcParams['legend.fontsize'] = 10
+    fig = plt.figure()
+    ax = fig.gca(projection='3d')
+    ax.plot_trisurf(dim1, dim2, dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+    ax.set_xlabel(colone)
+    ax.set_ylabel(coltwo)
+    ax.set_zlabel(value)
+    return plt
+
+for i,err in enumerate(errors):
+    file = file_prefix + str(err) + filename
+    df = pd.read_csv(file)
+    df['error'] = abs(df['Measured']-df['Expected'])
+    for m,e in zip(df['Measured'], df['error']):
+       dim1.append(err)
+       dim2.append(m)
+       dim3.append(e)
+
+colone="Fixation Error"
+coltwo="Measured Duration" 
+value="Duration Error"
+
+myplot = generate_2d_surface_plot(dim1, dim2, dim3, colone, coltwo, value)
+
+myplot.savefig("results/Error_surface.png")
+

### 2022-06-29 scripts/plot_measured_vs_expected.py

@@ -1,7 +1,7 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 
-file = "results/MREC_MAE_50/expected_values.csv"
+file = "results/MREC_MAE_60/expected_values.csv"
 df = pd.read_csv(file)
 
 measured = df['Measured']

### 2022-06-29 scripts/ANALYSE_RESULTS.sh
Adding final plot

@@ -8,3 +8,7 @@ python scripts/plot_measured_vs_expected.py
 # Generate the fixation versus duration error
 python scripts/plot_expected_errors.py
 
+# Error surface over MAE and Measured Duration
+python scripts/plot_error_surface.py
+
+

### 2022-06-29 scripts/plot_error_surface.py

@@ -8,8 +8,7 @@ import matplotlib as mpl
 file_prefix = "results/MREC_MAE_"
 filename = "/expected_values.csv"
 
-errors = [25,50,75,100,125,150]
-#errors = [20,40,60,80,100,120,140,160]
+errors = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
 
 D = 21
 dim1 = []

### 2022-06-29 paper/paper.tex
Paper cleanup

@@
 
-\title{Estimating Gaze Duration Error Distributions from Eye Tracking Data}
+\title{Estimating Gaze Duration Error from Eye Tracking Data}
 
@@
 a gaze duration measurement from an error in fixation points.
+
 In this work we develop an algorithm for estimating
 the error bounds on gaze duration measurement through monte carlo simulation based on the content
-of an eye tracking validation log file. We provide this algorithm as an open source application
+of an eye tracking calibration log file. We provide this algorithm as an open source application
 that allows for robust estimates of the error distribution of gaze duration measurements. 
@@
 
-We estimate the gaze duration measurement error through Monte Carlo simulation
-on the basis of the gaze fixation errors on the validation data from eye tracking
-fixation model. This error profile can be stratified into a grid across the screen dimensions. 
-For each of these grid locations we calculate two error probabilities, the false negative 
-probability (when true fixation is on the grid and the model produces coordinates outside the grid. 
-And the false positive probability, when the true fixation is outside the grid location and the model produces a prediction inside the grid square.
+We estimate the gaze duration measurement error through a Monte Carlo simulation
+using the eye tracking model's calibration file of fixation errors.
+The error profile of the eye tracking model allows us to generate sequences of true
+fixations and simulate measured fixations with the observed error of the model.
+These sequences of true and measured fixation are then converted into a distribution
+of expected gaze duration on a specifc area of interest, for a given measurement.
 
-We apply this grid of error probabilities in a Monte Carlo simulation that requires the following inputs:
-1.	The grid coordinates in which the item of concern is located.
-2.	The duration of true fixation on that item
-3.	The duration of the entire session of measurement.
-
-The Monte Carlo process will then repeatedly simulate sessions of the given length and gaze duration.
-Each simulation will use the error profiles to produce a potential model output.
-Collectively these estimates model outputs constitute a probability distribution over outputs for the given scenario.
+The algorithm consists of two core steps, first estimating the distrubtion of measurements
+for a given true duration. Secondly, inverting this into the distribution of true
+gaze durtions for a given measurement. Note, that the algorithm requires that the area of
+interest be defined as a fixed bounding box and that the distributions are estimated for
+a fixed session length. This session length is the total period of time that the area of
+interest was in a subject's field of view, and thus represents the maximum possible gaze
+duration.
 
@@
 For each point $f$ in $\mathcal{F}$ we determine a measurement point $\hat{f}$ 
-by drawing samples from the eye tracking validation set. 
-These points determine the $\delta_x$ and $\delta_y$ error in the measurement
-of gaze fixation point $f$. We draw these coordinates from the Calibration file such the
-the probability of the point being chosen is propotional to the distance of its target
-$x$ and $y$ from the coordinates of the point $f$ in the generated path. The complete algorithm
+by drawing samples from the eye tracking calibration set. 
+The samples from the calibration data are used to determine the $\delta_x$ and $\delta_y$ 
+error in the measurement of gaze fixation point $f$. 
+We draw these delta values from the calibration file such the
+the probability of the point being chosen is propotional to its distance from the point $f$
+in the generated path. The complete algorithm
 for estimating $P(\hat{d}|d,l,s)$ is shown in Algorithm \ref{alg:p_of_dhat}
@@
       \State $p \gets generatePath(d,l,s)$
-      \State $d_e \gets generateMeasurement(p,C)$  
+      \State $d_e \gets generateMeasurement(p,l,C)$  
       \State $P[d,d_e] += 1/N$          \Comment{Probability increment for a measuremnt of $d_e$}
@@
 true duration and the second the estimated duration. Note that our probability distribution is discrete
-because we restrict ourselves to sampling every increment point between zero and the session length.
+because we restrict ourselves to sampling over a fixed set of increments 
+between zero and the session length.
 
@@
 
+\subsection{Implementation}
+
+The algorithm described in the preceding sections has been implemented as an open source python 
+package called gazerr. It can be used as a software library, or deployed as a command line application. Source
+code is available on GitHub () and the package can be installed from PyPi.
+
+The source repository for the gazerr application also contains a series of scripts for generating data
+sets and running the experiments outlined in the next section. Full details available in the package
+README file.
+
 \subsection{Experiments}
@@
 and expected real duration under a variety of changing scenarios. These scenarios are variations
-in the underlying error indicated by the eye tracking validation file and length of the session
+in the underlying error indicated by the eye tracking calibration file and length of the session
 in which the user could be looking at the area of interest.
@@
 values (determined to be the mean Euclidean distance between the measured fixation and actual fixation).
-For each of these mean error values we generate a synthetic eye tracking validation file.
+For each of these mean error values we generate a synthetic eye tracking calibration file.
 
-We feed each of these synthetic validation files into gazerr to calculate posterior probability
+We feed each of these synthetic calibration files into gazerr to calculate posterior probability
 distributions over true gaze duration for the set of measured gaze durations. These posterior
@@
 session length on the expected gaze duration for a fixed set of measurements.
-
-All source code for these experiments is available in the gazerr open source package. 
-Raw data is generated by running the $RUN\_EXPERIMENTS.sh$ bash script. These results
-are then analysed and used to generatedt the resulting plots with the script 
-$ANALYSE\_RESULTS.sh$.
 
@@
 gaze duration and the mean error of fixation on the expected error in gaze duration.
+We display this as a three dimensional expected error surface over these two key dimensions
+in Figure \ref{fig:error_surface}
+
+\begin{figure}[h!]
+\includegraphics[scale=0.5]{../results/Error_surface.png}
+\caption{Mean gaze duration error over measured error and fixation MAE.}
+\label{fig:error_surface}
+\end{figure}
+
+There is a region in Figure \ref{fig:error_surface} that is unpopulated due to the fact that
+the estimated distributions contained no density in these regions. This was observed for 
+eye tracking models with high fixation error and when the measured duration was high. Meaning
+that our model estimates that when fixation error is this high, measurements of these values
+become extremely unlikley to be observed.
 

### 2022-06-29 scripts/RUN_BIAS_EXPERIMENTS.sh
Refining experiments and paper

@@ -1,0 +1,49 @@
+#!/bin/bash
+
+mkdir results/MREC_MAE_Biased_70
+python ./gazerr-runner.py data/validation_70_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_70
+
+mkdir results/MREC_MAE_Biased_20
+python ./gazerr-runner.py data/validation_20_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_20
+
+mkdir results/MREC_MAE_Biased_30
+python ./gazerr-runner.py data/validation_30_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_30
+
+mkdir results/MREC_MAE_Biased_40
+python ./gazerr-runner.py data/validation_40_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_40
+
+mkdir results/MREC_MAE_Biased_50
+python ./gazerr-runner.py data/validation_50_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_50
+
+mkdir results/MREC_MAE_Biased_60
+python ./gazerr-runner.py data/validation_60_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_60
+
+
+mkdir results/MREC_MAE_Biased_80
+python ./gazerr-runner.py data/validation_80_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_80
+
+mkdir results/MREC_MAE_Biased_90
+python ./gazerr-runner.py data/validation_90_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_90
+
+mkdir results/MREC_MAE_Biased_100
+python ./gazerr-runner.py data/validation_100_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_100
+
+mkdir results/MREC_MAE_Biased_110
+python ./gazerr-runner.py data/validation_110_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_110
+
+mkdir results/MREC_MAE_Biased_120
+python ./gazerr-runner.py data/validation_120_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_120
+
+mkdir results/MREC_MAE_Biased_130
[…]

### 2022-06-29 scripts/generate_datasets.py

@@ -14,11 +14,14 @@ device_height = 667
 
 N = 1500
 
-def add_noise(x,y,maerr):
+def add_noise(x,y,maerr, bias=False):
     factor = 1.85
     part = math.sqrt(math.pow(maerr,2)/2)
     x_err = random.uniform(-part*factor,part*factor)
     y_err = random.uniform(-part*factor,part*factor)
+    if bias:
+        x_err = -abs(x_err)
+        y_err = -abs(y_err)
     return x+x_err,y+y_err
 
 def euclidean(row):
@@ -29,7 +32,9 @@ def euclidean(row):
 
 for mae in maes:
     results = pd.DataFrame(columns=header)
+    results2 = pd.DataFrame(columns=header)
     filename = "data/validation_%s_MAE.csv"%str(mae)
+    filename2 = "data/validation_%s_MAE_Biased.csv"%str(mae)
     for n in range(N):
         my_x = random.randrange(0, device_width, 20)
         my_y = random.randrange(0, device_height, 20)
@@ -41,10 +46,19 @@ for mae in maes:
               "gaze_y":gaze_y
            }, ignore_index=True
         )
-    results.to_csv(filename, index=False,header=True)
-    results['error'] = results.apply(euclidean, axis=1)
+        gaze_x, gaze_y = add_noise(my_x,my_y,mae,bias=True)
+        results2 = results2.append({
+              "target_x":my_x,
+              "target_y":my_y,
+              "gaze_x":gaze_x,
+              "gaze_y":gaze_y
+           }, ignore_index=True
+        )
+    #results.to_csv(filename, index=False,header=True)
+    results2.to_csv(filename2, index=False,header=True)
+    results2['error'] = results.apply(euclidean, axis=1)
     print("Expected MAE: ", mae)
-    print("Simulated MAE: ", results['error'].mean())
+    print("Simulated MAE: ", results2['error'].mean())
 
 
 

### 2022-06-29 scripts/plot_error_surface.py

@@ -30,15 +30,17 @@ for i,err in enumerate(errors):
     df = pd.read_csv(file)
     df['error'] = abs(df['Measured']-df['Expected'])
     for m,e in zip(df['Measured'], df['error']):
-       dim1.append(err)
-       dim2.append(m)
-       dim3.append(e)
+       if m!=e:
+          dim1.append(err)
+          dim2.append(m)
+          dim3.append(e)
 
 colone="Fixation Error"
 coltwo="Measured Duration" 
 value="Duration Error"
 
 myplot = generate_2d_surface_plot(dim1, dim2, dim3, colone, coltwo, value)
+myplot.show()
 
-myplot.savefig("results/Error_surface.png")
+#myplot.savefig("results/Error_surface.png")
 

### 2022-06-29 scripts/plot_error_surfaces.py

@@ -1,0 +1,82 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import expon
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import matplotlib as mpl
+
+file_prefix = "results/MREC_MAE_"
+filename = "/expected_values.csv"
+
+errors = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
+
+D = 21
+dim1 = []
+dim2 = []
+dim3 = []
+
+v2_dim1 = []
+v2_dim2 = []
+v2_dim3 = []
+
+def generate_2d_surface_plot(ax, dim1, dim2, dim3, colone, coltwo, value):
+    ax.plot_trisurf(dim1, dim2, dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+    ax.set_xlabel(colone)
+    ax.set_ylabel(coltwo)
+    ax.set_zlabel(value)
+    return plt
+
+for i,err in enumerate(errors):
+    file = file_prefix + str(err) + filename
+    df = pd.read_csv(file)
+    df['error'] = abs(df['Measured']-df['Expected'])
+    for m,e in zip(df['Measured'], df['error']):
+       dim1.append(err)
+       dim2.append(m)
+       dim3.append(e)
+       if m!=e:
+          v2_dim1.append(err)
+          v2_dim2.append(m)
+          v2_dim3.append(e)
+
+colone="Fixation Error"
+coltwo="Measured Duration" 
+value="Duration Error"
+
+# set up a figure twice as wide as it is tall
+fig = plt.figure(figsize=plt.figaspect(0.5))
+
+# =============
+# First subplot
+# =============
+# set up the axes for the first plot
+ax = fig.add_subplot(1, 2, 1, projection='3d')
+
+# plot a 3D surface like in the example mplot3d/surface3d_demo
+
+#surf = ax.plot_surface(dim1, dim2, dim3, rstride=1, cstride=1, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+ax.plot_trisurf(dim1, dim2, dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+ax.set_xlabel(colone)
+ax.set_ylabel(coltwo)
+ax.set_zlabel("")
+#fig.colorbar(surf, shrink=0.5, aspect=10)
+
+# ==============
+# Second subplot
+# ==============
[…]

### 2022-06-29 scripts/plot_expected_errors.py

@@ -33,12 +33,12 @@ fig = plt.figure(figsize=(10,10))
 ax = fig.add_subplot()
 
 ax.set_title('Fixation Vs Duration Expected Error')
-ax.plot(errors, gaze_err_uni, c='darkorange', label="Uniform")
-ax.plot(errors, gaze_err_exp, c='teal', label="Exponential")
+ax.plot(errors, gaze_err_uni, c='darkorange', label="Uniform Measurement Distribution")
+ax.plot(errors, gaze_err_exp, c='teal', label="Exponential Measurement Distribution")
 ax.set_xlabel('Gaze Fixation μ Error (px)', fontsize=15)
 ax.set_ylabel('Gaze Duration μ Error (ms)', fontsize=15)
 ax.set_xlim([0, 175])
-ax.legend()
+ax.legend( prop={'size': 15} )
 
 plt.savefig("results/Fixation_vs_duration_error.png")
 

### 2022-06-29 scripts/plot_measured_vs_expected.py

@@ -1,21 +1,30 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 
-file = "results/MREC_MAE_60/expected_values.csv"
+file = "results/MREC_MAE_70/expected_values.csv"
+file2 = "results/MREC_MAE_Biased_70/expected_values.csv"
+
 df = pd.read_csv(file)
+df2 = pd.read_csv(file2)
 
 measured = df['Measured']
 expected = df['Expected']
+measured2 = df2['Measured']
+expected2 = df2['Expected']
 
-plt.figure(figsize=(10,10))
-plt.scatter(measured, expected, c='crimson')
+fig = plt.figure(figsize=(10,10))
 
+ax = fig.add_subplot()
+ax.set_title('Measured Duration Vs Expected Duration')
+ax.plot(measured, expected, c='orchid', label="Uniform Fixation Error")
+ax.plot(measured2, expected2, c='teal', label="Biased Fixation Error", linestyle='dashed')
 p1 = max(max(measured), max(expected))
 p2 = min(min(measured), min(expected))
-plt.plot([p1, p2], [p1, p2], 'b-')
-plt.xlabel('Measured Gaze Duration', fontsize=15)
-plt.ylabel('Expected Gaze Duration', fontsize=15)
-plt.axis('equal')
+ax.plot([p1, p2], [p1, p2], c='lightsteelblue', linestyle='dotted', label='Ideal' )
+ax.set_xlabel('Measured Gaze Duration', fontsize=15)
+ax.set_ylabel('Expected Gaze Duration', fontsize=15)
+ax.legend( prop={'size': 15} )
 
 plt.savefig("results/Measured_vs_expected.png")
+
 

### 2022-06-30 scripts/plot_error_surfaces.py
Improving experimental results, plotting and paper

@@ -44,39 +44,31 @@ coltwo="Measured Duration"
 value="Duration Error"
 
 # set up a figure twice as wide as it is tall
-fig = plt.figure(figsize=plt.figaspect(0.5))
+#fig = plt.figure(figsize=plt.figaspect(0.5))
+fig = plt.figure(figsize=(18,8))
 
 # =============
 # First subplot
 # =============
-# set up the axes for the first plot
 ax = fig.add_subplot(1, 2, 1, projection='3d')
-
-# plot a 3D surface like in the example mplot3d/surface3d_demo
-
-#surf = ax.plot_surface(dim1, dim2, dim3, rstride=1, cstride=1, cmap=cm.coolwarm, linewidth=0, antialiased=False)
-ax.plot_trisurf(dim1, dim2, dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+surf = ax.plot_trisurf(dim1, dim2, dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
 ax.set_xlabel(colone)
 ax.set_ylabel(coltwo)
 ax.set_zlabel("")
-#fig.colorbar(surf, shrink=0.5, aspect=10)
+ax.set_title('A) Error Surface - All Measurements', y=-0.1)
 
 # ==============
 # Second subplot
 # ==============
 # set up the axes for the second plot
 ax = fig.add_subplot(1, 2, 2, projection='3d')
-
-# plot a 3D wireframe like in the example mplot3d/wire3d_demo
-#surf = ax.plot_surface(v2_dim1, v2_dim2, v2_dim3, rstride=1, cstride=1, cmap=cm.coolwarm,
-#                       linewidth=0, antialiased=False)
-ax.plot_trisurf(v2_dim1, v2_dim2, v2_dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+surf = ax.plot_trisurf(v2_dim1, v2_dim2, v2_dim3, cmap=cm.coolwarm, linewidth=0, antialiased=False)
 ax.set_xlabel(colone)
 ax.set_ylabel(coltwo)
 ax.set_zlabel(value)
-
+fig.colorbar(surf, shrink=0.5, aspect=10)
+ax.set_title('B) Error Surface - Excluding No Density Regions', y=-0.26)
 plt.show()
-
 
 #plt.savefig("results/Error_surfaces.png")
 

### 2022-06-30 scripts/plot_expected_errors.py

@@ -4,11 +4,14 @@ from scipy.stats import expon
 import matplotlib.pyplot as plt
 
 file_prefix = "results/MREC_MAE_"
+file_prefix2 = "results/MREC_MAE_Biased_"
 filename = "/expected_values.csv"
 
 errors = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
 gaze_err_uni = []
 gaze_err_exp = []
+gaze_err2_uni = []
+gaze_err2_exp = []
 
 # This creates the weighting for measurements
 x = expon.rvs(scale=30, size=10000)
@@ -19,26 +22,50 @@ dist_uni = [1/len(vallist) for x in vallist]
 
 for err in errors:
     file = file_prefix + str(err) + filename
+    file2 = file_prefix2 + str(err) + filename
     df = pd.read_csv(file)
+    df2 = pd.read_csv(file2)
     df['error'] = abs(df['Measured']-df['Expected'])
+    df2['error'] = abs(df2['Measured']-df2['Expected'])
     df['uni'] = dist_uni
     df['exp'] = dist_exp
     df['err_uni'] = df['error']*df['uni'] 
     df['err_exp'] = df['error']*df['exp'] 
     gaze_err_uni.append(df['err_uni'].sum())
     gaze_err_exp.append(df['err_exp'].sum()) 
+    df2['uni'] = dist_uni
+    df2['exp'] = dist_exp
+    df2['err_uni'] = df2['error']*df2['uni'] 
+    df2['err_exp'] = df2['error']*df2['exp'] 
+    gaze_err2_uni.append(df2['err_uni'].sum())
+    gaze_err2_exp.append(df2['err_exp'].sum()) 
 
-fig = plt.figure(figsize=(10,10))
+fig = plt.figure(figsize=(18,8))
+#fig = plt.figure(figsize=plt.figaspect(0.5))
 
-ax = fig.add_subplot()
-
-ax.set_title('Fixation Vs Duration Expected Error')
+# =============
+# First subplot
+# =============
+# set up the axes for the first plot
+ax = fig.add_subplot(1, 2, 1)
+ax.set_title('A) Expected Error - Gaze Fixation Vs Duration')
 ax.plot(errors, gaze_err_uni, c='darkorange', label="Uniform Measurement Distribution")
[…]

### 2022-06-30 scripts/plot_measured_vs_expected.py

@@ -12,7 +12,7 @@ expected = df['Expected']
 measured2 = df2['Measured']
 expected2 = df2['Expected']
 
-fig = plt.figure(figsize=(10,10))
+fig = plt.figure(figsize=(12,6))
 
 ax = fig.add_subplot()
 ax.set_title('Measured Duration Vs Expected Duration')

### 2022-07-03 paper/paper.tex
Refining experiments and paper
Improving experimental results, plotting and paper
Finalising paper draft

@@
 values (determined to be the mean Euclidean distance between the measured fixation and actual fixation).
-For each of these mean error values we generate a synthetic eye tracking calibration file.
+For each of these mean error values we generate a synthetic eye tracking calibration file. We produce
+two versions of this calibration data, one in which the error is distributed around the true target,
+and a second biased version in which the error tends toward the upper left of the true position. We 
+include the biased calibration data as it is consistent with our observations of real eye tracking data
+and we explore the extent to which this bias affects gaze duration measurements.
 
@@
 multiple variations of the simulation configuration. At lower meaurement values there is a 
-tendency for the measurement to be an over estimation, and at higher values an under-estimation.
+tendency for the measurement to be an accurate estimation, while at higher values it tends
+toward an under-estimation of the true gaze duration. Surprisingly the results using the
+biased calibration data seem to mitigate this effect somewhat, but the overall pattern 
+remains consistent.
 
@@
 \begin{figure}[h!]
-\includegraphics[scale=0.5]{../results/Fixation_vs_duration_error.png}
+\centering
+\includegraphics[scale=0.4]{../results/Fixation_vs_duration_error.png}
 \caption{Fixation error versus gaze duration error.}
@@
 grows with error in fixation points, as would be expected. However, when we use the exponential
-weighting to calculate the expected error in duration, the gaze duration error grows much slower
+weighting to calculate the expected error in duration the gaze duration error grows much slower
 than in the uniform case. This suggests that the observed tendency toward smaller measurements
-in area of interest studies means lower expected error.
+in area of interest studies means lower expected error overall. This pattern is true for the 
+centrally distributed error (A) as well as the biased calibration error (B).
 
@@
 \begin{figure}[h!]
-\includegraphics[scale=0.5]{../results/Error_surface.png}
+\centering
+\includegraphics[scale=0.4]{../results/Error_surface.png}
 \caption{Mean gaze duration error over measured error and fixation MAE.}
@@
 
-There is a region in Figure \ref{fig:error_surface} that is unpopulated due to the fact that
-the estimated distributions contained no density in these regions. This was observed for 
+The left hand plot shows the complete error surface in which you will observe large spikes in
+error in the corner correspinding to high MAE for the fixation models and high measurements of
+gaze duration. This effect is caused by a complete absence of probability density in these
+regions, meaning that the error is maximal because the expected value is zero. This effect 
+was consistenlt observed for 
 eye tracking models with high fixation error and when the measured duration was high. Meaning
@@
 become extremely unlikley to be observed.
+
+We added the right hand plot in Figure \ref{fig:error_surface} to illustrate the error surface
+in the absence of these deceptive outliers. We see that in general the expected error in a gaze
+duration is maximal around the mid point of the measurement range. The effect is lessened by 
+lower MAE in the fixation models, but remains consistent for all levels of fixation error.
 

### 2022-07-18 paper/paper.tex
Paper edits

@@
 The algorithm described in the preceding sections has been implemented as an open source python 
-package called gazerr. It can be used as a software library, or deployed as a command line application. Source
-code is available on GitHub () and the package can be installed from PyPi.
+package called \textit{gazerr}. 
+It can be used as a software library, exposing functions for calculating the probability densities,
+or it can deployed as a command line application. 
+The source code is available on GitHub and the python package can be installed from PyPi.
 
@@
 data can be utilised for estimating errors in gaze duration estimation. The process involves
-simulating the distribution of measurements for a given true durastion, and then inverting
-it into a distribution over true duration for a given measurement.
+simulating the distribution of measured duration for each possible true duration, 
+and then inverting it with Bayes rule into a distribution over true duration for a given measurement.
 
-We have released this approach as the open source application gazerr, and used it to explore
+We have released this approach as the open source application \textit{gazerr} and used it to explore
 the relationship between fixation error and duration error. Our results show that the nature
 of the error in gaze duration depends on the size of the measurement. Smaller measurements
-tend toward over prediction, and larger measurements towards under prediction. We plotted
+tend to have lower expected error while larger measurements tend toward under prediction. We plotted
 the relationship between expected error and saw that observed tendency toward smaller duration
 measurements (estimated with an psuedo-exponential distribution) delivers a slower growing
-expected duraiton error with the mean error in fixation points.
+expected duration error with the mean error in fixation points.
 
 Finally, our simualtions revealed that for large error in fixation points certain measurements
-become some unlikely that our monte carlo simualtions placed none of the probability density
+become so unlikely that our monte carlo simualtions placed none of the probability density
 in those regions. This strongly suggests that an additional downside to large fixation point error

### 2022-07-24 scripts/generate_datasets.py
dataset generator

@@ -54,7 +54,7 @@ for mae in maes:
               "gaze_y":gaze_y
            }, ignore_index=True
         )
-    #results.to_csv(filename, index=False,header=True)
+    results.to_csv(filename, index=False,header=True)
     results2.to_csv(filename2, index=False,header=True)
     results2['error'] = results.apply(euclidean, axis=1)
     print("Expected MAE: ", mae)

### 2022-07-24 paper/paper.tex
paper

@@
 In this work we develop an algorithm for estimating
-the error bounds on gaze duration measurement through monte carlo simulation based on the content
-of an eye tracking calibration log file. We provide this algorithm as an open source application
-that allows for robust estimates of the error distribution of gaze duration measurements. 
+the error distribution of gaze duration measurement through monte carlo simulation 
+using the content of an eye tracking calibration log file. 
+We provide this algorithm in an open source application
+to allow researchers to understand the error in their gaze duration measurements. 
 We use this application to conduct experiments on the expected error bounds for different duration
-measurements across a range of session lengths and area of interest positions within the viewport
-of mobile devices. The results indicate that error in gaze duration estimation is sensitive to
-the size of fixation error as well as the proportion of total field of view that the object occupies
-and the proportion of viewing time that was taken with gaze fixation.
+measurements across a fixed session length for a simulated advertising area of interest 
+study on a mobile device. 
+The results indicate that error in gaze duration estimation is sensitive to
+fixation error beyond a bound that will depend on the size of interest and the session length.
 

### 2022-09-07 paper/paper.tex
paper and synth data changes

@@
 gaze fixation point measurements. These applications are often used to infer a gaze duration
-metric from a series of fixation measurements. There is no direct method for infering the error in
+metric from a series of fixation measurements. There is no direct method for inferring the error in
 a gaze duration measurement from an error in fixation points.
@@
 In this work we develop an algorithm for estimating
-the error distribution of gaze duration measurement through monte carlo simulation 
+the error distribution of gaze duration measurement through Monte Carlo simulation 
 using the content of an eye tracking calibration log file. 

### 2022-09-23 paper/paper.tex
Paper adjustments

@@
 Eye tracking applications produce a series of gaze fixation points that can be attributed to
-objects within a subject's field of vision. Error is typically measured on the basis of individual
-gaze fixation point measurements. These applications are often used to infer a gaze duration
-metric from a series of fixation measurements. There is no direct method for inferring the error in
+objects within a subject's field of vision. Error is typically measured on the basis 
+of individual gaze fixation point measurements. These applications are often used 
+to infer a gaze duration metric from a series of fixation measurements. 
+There is no direct method for inferring the error in
 a gaze duration measurement from an error in fixation points.
 
-In this work we develop an algorithm for estimating
-the error distribution of gaze duration measurement through Monte Carlo simulation 
-using the content of an eye tracking calibration log file. 
+In this work we develop an algorithm for estimating the error distribution of 
+gaze duration measurement through Monte Carlo simulation using the content 
+of an eye tracking calibration log file. 
 We provide this algorithm in an open source application
 to allow researchers to understand the error in their gaze duration measurements. 
-We use this application to conduct experiments on the expected error bounds for different duration
-measurements across a fixed session length for a simulated advertising area of interest 
-study on a mobile device. 
+We use this application to conduct experiments on the expected error bounds for 
+different duration measurements across a fixed session length for a simulated 
+advertising area of interest study on a mobile device. 
 The results indicate that error in gaze duration estimation is sensitive to
@@
 
-Eye tracking is an important technology with a wide variety of applications. It permits evaluation
-of software user interfaces\cite{Harezlak2015}, the development of new forms of software interaction, as well as a potential method of biometric identification\cite{Kasprowski2018}. 
-Eye tracking has become a core tool for empirical investigations 
-into human behaviour and is widely used as a method of measuring explicit attention to visual stimuli. 
+Eye tracking is an important technology with a wide variety of applications. 
+It permits evaluation of software user interfaces\cite{Harezlak2015}, 
+the development of new forms of software interaction, as well as a potential method of biometric identification\cite{Kasprowski2018}. Eye tracking has become a core tool for empirical 
+investigations into human behaviour and is widely used as a method of measuring explicit 
+attention to visual stimuli. 
 It has been applied to study psychological phenomena ranging from cognition\cite{Brunye2019}
@@
 are an ongoing focus for the development of algorithms\cite{Zhang2014,Hassoumi2019} 
-and software tools\cite{ETCAL2018}
+and software tools\cite{ETCAL2018}. In some scenarios there are additional considerations
+such as whether to use fixation data from individual eyes, or a composite 
+signal\cite{Hooge2019}.
 

### 2022-09-23 paper/paper.tex
paper update

@@
 
-Increasingly, eye tracking applications are built using machine learning models trained on 
-data sets containing combinations of facial images and fixation coordinates collected 
-through careful application design. The resulting gaze fixation prediction model can 
-then be applied sequentially to determine how long a subject fixated on a given object.
-Thus, gaze duration is derived from aggregation of gaze fixation.
+Increasingly, eye tracking applications are built using devices that employ 
+machine learning models trained to predict fixation locations from a camera image.
+While this approach to eye tracking may suffer in terms of accuracy, it can
+open up possibilities to researchers that head position restricting devices prevent
+\cite{Valtakari2021}.
 
@@
 signal\cite{Hooge2019}.
+
+Discuss error elments around accuracy and precision \cite{Holmqvist2022}
 

### 2022-09-26 paper/paper.tex
Adding refs and lit review

@@
 The errors in fixation point measurement for a given eye tracking solution are routinely
-adjusted through a process called calibration. Improvements in the callibration process
+monitored and adjusted through a process called calibration, in which users are presented
+a series or \it{required fixation locations} (RFL) which are compared with measured 
+points of fixation\cite{Hornof2002}. 
+Improvements in the callibration process
 are an ongoing focus for the development of algorithms\cite{Zhang2014,Hassoumi2019} 
@@
 
-Discuss error elments around accuracy and precision \cite{Holmqvist2022}
+Typically the nature of the error detected in the calibration process can be broken
+into both accuracy and precisions. Wherein accuracy is the tendency for the measure
+fixation points to be closer to true fixation, and precision is the tendency for the
+measured locations to vary for a given true fixation\cite{Holmqvist2022}. This distinction
+is critical because it forms the basis of methods that can correct for the kind of 
+systematic bias observed with low precision eye tracking.
 

### 2022-10-12 paper/paper.tex
Paper update

@@
 
-Eye tracking applications produce a series of gaze fixation points that can be attributed to
-objects within a subject's field of vision. Error is typically measured on the basis 
-of individual gaze fixation point measurements. These applications are often used 
-to infer a gaze duration metric from a series of fixation measurements. 
-There is no direct method for inferring the error in
-a gaze duration measurement from an error in fixation points.
+Eye tracking technologies produce a series of timestamped gaze fixation points 
+that can be attributed to objects within a subject's field of vision. 
+These fixation points are often aggregated in order to generate a measurement
+of the gaze duration in Area of Interest studies.
+Error is typically measured on the basis of the variation between individual 
+gaze fixation points and the intended point of fixation.
+These underlying errors in fixation have no inherent transformation for making
+a statement about error in gaze duration, even though they are expected to be
+related.
 
@@
 different duration measurements across a fixed session length for a simulated 
-advertising area of interest study on a mobile device. 
+area of interest study on a mobile device. 
 The results indicate that error in gaze duration estimation is sensitive to
-fixation error beyond a bound that will depend on the size of interest and the session length.
+fixation error beyond a bound that will depend on the size of the area of interest 
+and the comparative length of the viewing session.
 
@@
 While this approach to eye tracking may suffer in terms of accuracy, it can
-open up possibilities to researchers that head position restricting devices prevent
-\cite{Valtakari2021}.
+open up possibilities for researchers in terms of reduced cost and increased scale
+of experiments. As well as facilitating studies that head position restricting 
+devices prevent\cite{Valtakari2021}.
 
@@
 in the training data. In addition, it has been observed that the error profile varies 
-depending on the location of the true fixation point within the subjects field of view.
+depending on the location of the true fixation point within the subject's field of view.
 The errors in fixation point measurement for a given eye tracking solution are routinely
 monitored and adjusted through a process called calibration, in which users are presented
-a series or \it{required fixation locations} (RFL) which are compared with measured 
+a series or \emph{required fixation locations} (RFL) which are compared with measured 
 points of fixation\cite{Hornof2002}. 
@@
 Typically the nature of the error detected in the calibration process can be broken
-into both accuracy and precisions. Wherein accuracy is the tendency for the measure
+into both accuracy and precision. Wherein accuracy is the tendency for the measure
 fixation points to be closer to true fixation, and precision is the tendency for the
@@
 is critical because it forms the basis of methods that can correct for the kind of 
-systematic bias observed with low precision eye tracking.
+systematic bias observed with low precision eye tracking. It also been observed that
+metrics of the precision in these models vary in the way they rank technologies\cite{Niehorster2020}. We can conclude from this that the patterns of systematic error in eye tracking
+data are more complicated that can be quanitifed by a single metric.
 
@@
 Influencing factors include the ratio of measured gaze duration to the total viewing time
-and the error profile of the machine learning fixation coordinate model.
+and the error profile of the gaze fixation technology.
 

### 2022-10-19 scripts/generate_datasets.py
Changing synthetic data creation to include high error high precision simulation

@@ -14,12 +14,14 @@ device_height = 667
 
 N = 1500
 
-def add_noise(x,y,maerr, bias=False):
+def add_noise(x,y,maerr, bias=False, precise=False):
     factor = 1.85
     part = math.sqrt(math.pow(maerr,2)/2)
     x_err = random.uniform(-part*factor,part*factor)
     y_err = random.uniform(-part*factor,part*factor)
     if bias:
+        if precise:
+            y_err = x_err + random.uniform(-0.35*x_err, 0.35*x_err)
         x_err = -abs(x_err)
         y_err = -abs(y_err)
     return x+x_err,y+y_err
@@ -33,8 +35,10 @@ def euclidean(row):
 for mae in maes:
     results = pd.DataFrame(columns=header)
     results2 = pd.DataFrame(columns=header)
-    filename = "data/validation_%s_MAE.csv"%str(mae)
-    filename2 = "data/validation_%s_MAE_Biased.csv"%str(mae)
+    results3 = pd.DataFrame(columns=header)
+    filename = "data/validation_%s_MErr.csv"%str(mae)
+    filename2 = "data/validation_%s_MErr_Biased.csv"%str(mae)
+    filename2 = "data/validation_%s_MErr_Biased_Precise.csv"%str(mae)
     for n in range(N):
         my_x = random.randrange(0, device_width, 20)
         my_y = random.randrange(0, device_height, 20)
@@ -54,11 +58,24 @@ for mae in maes:
               "gaze_y":gaze_y
            }, ignore_index=True
         )
+        gaze_x, gaze_y = add_noise(my_x,my_y,mae,bias=True,True)
+        results3 = results3.append({
+              "target_x":my_x,
+              "target_y":my_y,
+              "gaze_x":gaze_x,
+              "gaze_y":gaze_y
+           }, ignore_index=True
+        )
     results.to_csv(filename, index=False,header=True)
     results2.to_csv(filename2, index=False,header=True)
-    results2['error'] = results.apply(euclidean, axis=1)
-    print("Expected MAE: ", mae)
-    print("Simulated MAE: ", results2['error'].mean())
+    results3.to_csv(filename3, index=False,header=True)
+    results['error'] = results.apply(euclidean, axis=1)
+    results2['error'] = results2.apply(euclidean, axis=1)
+    results3['error'] = results3.apply(euclidean, axis=1)
+    print("Expected Error: ", mae)
[…]

### 2022-10-19 paper/paper.tex
paper updates

@@
 
-Gaze fixation models will typically have an error profile that is dependent on both 
-the nature of the media being presented, the variety of faces and lighting conditions 
-in the training data. In addition, it has been observed that the error profile varies 
-depending on the location of the true fixation point within the subject's field of view.
+The source of errors in eye tracking technology can be decomposed into a range of
+independent sources that researchers need to consider\cite{Holmqvist2012}.
+Unless they are properly addressed these sources of error can manifest themselves
+as systematic biases across undesirable dimensions such as the age\cite{Dalrymple2018}
+or ethnicity of subjects\cite{Blignaut2013}. Machine learning based eye tracking solutions
+can be impacted by the variety of faces and lighting conditions in the training data. 
 The errors in fixation point measurement for a given eye tracking solution are routinely
@@
 
-Typically the nature of the error detected in the calibration process can be broken
-into both accuracy and precision. Wherein accuracy is the tendency for the measure
-fixation points to be closer to true fixation, and precision is the tendency for the
-measured locations to vary for a given true fixation\cite{Holmqvist2022}. This distinction
+In eye tracking literature the error detected in the calibration process is classified
+using two independent metrics; accuracy and precision. Accuracy is a measure for how
+close, on average, the fixation points are to the true point of fixation.
+Precision is the tendency for the measured locations to vary between subsequent measurements
+for a given point of fixation\cite{Holmqvist2022}. This distinction
 is critical because it forms the basis of methods that can correct for the kind of 
 systematic bias observed with low precision eye tracking. It also been observed that
-metrics of the precision in these models vary in the way they rank technologies\cite{Niehorster2020}. We can conclude from this that the patterns of systematic error in eye tracking
-data are more complicated that can be quanitifed by a single metric.
+metrics of the precision in these models vary in the way they rank 
+technologies\cite{Niehorster2020}. We can conclude from this that the patterns of 
+systematic error in eye tracking
+data are more complicated that can be quantifed by a single metric.
 
@@
 specifics of the model and the conditions under which it is being used.
-Influencing factors include the ratio of measured gaze duration to the total viewing time
-and the error profile of the gaze fixation technology.
-
-The errors of eye tracking data technology can be decomposed into a range of
-independent sources that researchers need to consider \cite{Holmqvist2012}.
-Unless they are properly addressed these sources of error can manifest themselves
-as systematic biases across undesirable dimensions such as the age\cite{Dalrymple2018}
-or ethnicity of subjects\cite{Blignaut2013}.
+Influencing factors include the ratio of measured gaze duration to the 
+total viewing time and the error profile of the gaze fixation technology.
 
@@
 by drawing samples from the eye tracking calibration set.
-The samples from the calibration data are used to determine the $\delta_x$ and $\delta_y$ 
-error in the measurement of gaze fixation point $f$. 
+
+The samples from the calibration data are drawn as a delta in the two dimensional Cartesian
+space of the field of view. Each delta consists of $\delta_x$ and $\delta_y$ which correspond
+to the error in the $x$ and $y$ dimensions. We apply these delta values to the true gaze 
+fixation point $f$ to create the simulated measurement $\hat{f}$.
 We draw these delta values from the calibration file such the
-the probability of the point being chosen is propotional to its distance from the point $f$
-in the generated path. The complete algorithm
-for estimating $P(\hat{d}|d,l,s)$ is shown in Algorithm \ref{alg:p_of_dhat}
+the probability of choosing a delta is propotional to the distance between the point $f$
+in the generated path, and the true fixation point in the callibration set.
+
+The complete algorithm for estimating $P(\hat{d}|d,l,s)$ is shown in Algorithm \ref{alg:p_of_dhat}
 
@@
 true duration and the second the estimated duration. Note that our probability distribution is discrete
-because we restrict ourselves to sampling over a fixed set of increments 
+because we restrict ourselves to sampling over a set of fixed time period increments ($i$) 
 between zero and the session length.
@@
 estimate of gaze duration. Meaning that the probability distribution we want is the distribution 
-over true gaze durations for a given measurement, rather than the inverse which is what we 
-have produced in the previous section. We produce an estimate of $P(d|\hat{d},l,s)$
-through an application of Bayes' rule, as shown in Equation \ref{eq:p_of_d}.
+over true gaze durations for a given measurement. In the previous section we estimated the inverse of
+this: the distribution of measurements of a given true duration.
+
+In order to estimate the desired distribution $P(d|\hat{d},l,s)$ we need to use 
+Bayes' rule, as shown in Equation \ref{eq:p_of_d}.
 
@@
 discrete estimations of an underlying continuous distribution the value of $P(d)$ is
-equal to $1/(1+floor(10*s))$
- We can estimate the value
-of the denominator $P(\hat{d},l,s)$ by iterating over all values of d and summing the product
+equal to $1/(1+floor(i*s))$. We can estimate the value
+of the denominator $P(\hat{d},l,s)$ by iterating over all values of $d$ and summing the product
 of $ P(\hat{d},l,s|d) \dot P(d)$. This fully explicated form is shown in 
@@
 
-For the sake of our simulations we use a fixed sized simulated device with viewport pixel 
+For the sake of our simulations we use a simulated device with viewport pixel 
 dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium rectangle ad unit (MREC)
 that is located toward the top of the screen. We iterate over a range of potential mean error
-values (determined to be the mean Euclidean distance between the measured fixation and actual fixation).
+values (equivalent to varying the accuracy), determined to by the mean Euclidean distance 
+between the measured fixation point and actual fixation points.
 For each of these mean error values we generate a synthetic eye tracking calibration file. We produce
-two versions of this calibration data, one in which the error is distributed around the true target,
+two versions of this calibration data, one in which the error is centered around the true target,
 and a second biased version in which the error tends toward the upper left of the true position. We 
@@
 The left hand plot shows the complete error surface in which you will observe large spikes in
-error in the corner correspinding to high MAE for the fixation models and high measurements of
-gaze duration. This effect is caused by a complete absence of probability density in these
+error in the corner corresponding to both high mean error for the fixation models and 
+high measurements of gaze duration. 
+This effect is caused by a complete absence of probability density in these
 regions, meaning that the error is maximal because the expected value is zero. This effect 
-was consistenlt observed for 
-eye tracking models with high fixation error and when the measured duration was high. Meaning
-that our model estimates that when fixation error is this high, measurements of these values
+was consistently observed for 
+eye tracking models with high mean fixation error and when the measured duration was high. 
+The consequence of this is that our model estimates that when fixation error is this high, 
+measurements of these values
 become extremely unlikley to be observed.
@@
 We added the right hand plot in Figure \ref{fig:error_surface} to illustrate the error surface
-in the absence of these deceptive outliers. We see that in general the expected error in a gaze
-duration is maximal around the mid point of the measurement range. The effect is lessened by 
-lower MAE in the fixation models, but remains consistent for all levels of fixation error.
+in the absence of these deceptive outliers. We see that, in general, the expected error in a gaze
+duration measurement is maximal around the mid point of the measurement range. The effect is lessened by 
+lower mean error in the fixation models, but remains consistent for all levels of fixation error.
 
@@
 
-We have demonstrated that the statistial information contained in eye tracking validation
+We have demonstrated that the statistial information contained in eye tracking calibration
 data can be utilised for estimating errors in gaze duration estimation. The process involves
@@
 
-Finally, our simualtions revealed that for large error in fixation points certain measurements
-become so unlikely that our monte carlo simualtions placed none of the probability density
+Finally, our simulations revealed that for large error in fixation points, certain measurements
+become so unlikely that our Monte Carlo simualtions placed none of the probability density
 in those regions. This strongly suggests that an additional downside to large fixation point error

### 2022-10-20 scripts/ANALYSE_RESULTS.sh
Changing experiments and analysis to include biased precise experiments

@@ -1,6 +1,10 @@
 #!/bin/bash
 
-# NOTE: This script depends on the output of RUN_EXPERIMENTS.sh
+# NOTE: This script depends on the outputs of the folling scripts:
+# RUN_EXPERIMENTS.sh
+# RUN_BIAS_EXPERIMENTS.sh
+# RUN_PRECISE_BIAS_EXPERIMENTS.sh
+#
 
 # Generate the Measured Versus Expected Plot
 python scripts/plot_measured_vs_expected.py

### 2022-10-20 scripts/RUN_BIAS_EXPERIMENTS.sh

@@ -1,49 +1,49 @@
 #!/bin/bash
 
-mkdir results/MREC_MAE_Biased_70
-python ./gazerr-runner.py data/validation_70_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_70
 
-mkdir results/MREC_MAE_Biased_20
-python ./gazerr-runner.py data/validation_20_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_20
+mkdir results/MREC_MErr_Biased_20
+python ./gazerr-runner.py data/validation_20_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_20
 
-mkdir results/MREC_MAE_Biased_30
-python ./gazerr-runner.py data/validation_30_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_30
+mkdir results/MREC_MErr_Biased_30
+python ./gazerr-runner.py data/validation_30_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_30
 
-mkdir results/MREC_MAE_Biased_40
-python ./gazerr-runner.py data/validation_40_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_40
+mkdir results/MREC_MErr_Biased_40
+python ./gazerr-runner.py data/validation_40_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_40
 
-mkdir results/MREC_MAE_Biased_50
-python ./gazerr-runner.py data/validation_50_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_50
+mkdir results/MREC_MErr_Biased_50
+python ./gazerr-runner.py data/validation_50_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_50
 
-mkdir results/MREC_MAE_Biased_60
-python ./gazerr-runner.py data/validation_60_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MAE_Biased_60
+mkdir results/MREC_MErr_Biased_60
+python ./gazerr-runner.py data/validation_60_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_60
 
+mkdir results/MREC_MErr_Biased_70
[…]

### 2022-10-20 scripts/RUN_EXPERIMENTS.sh

@@ -1,48 +1,48 @@
 #!/bin/bash
 
-mkdir results/MREC_MAE_20
-python ./gazerr-runner.py data/validation_20_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_20
+mkdir results/MREC_MErr_20
+python ./gazerr-runner.py data/validation_20_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_20
 
-mkdir results/MREC_MAE_30
-python ./gazerr-runner.py data/validation_30_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_30
+mkdir results/MREC_MErr_30
+python ./gazerr-runner.py data/validation_30_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_30
 
-mkdir results/MREC_MAE_40
-python ./gazerr-runner.py data/validation_40_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_40
+mkdir results/MREC_MErr_40
+python ./gazerr-runner.py data/validation_40_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_40
 
-mkdir results/MREC_MAE_50
-python ./gazerr-runner.py data/validation_50_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_50
+mkdir results/MREC_MErr_50
+python ./gazerr-runner.py data/validation_50_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_50
 
-mkdir results/MREC_MAE_60
-python ./gazerr-runner.py data/validation_60_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_60
+mkdir results/MREC_MErr_60
+python ./gazerr-runner.py data/validation_60_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_60
 
-mkdir results/MREC_MAE_70
-python ./gazerr-runner.py data/validation_70_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_70
+mkdir results/MREC_MErr_70
+python ./gazerr-runner.py data/validation_70_MErr.csv 400 1000 40,40 340,290 results/MREC_MErr_70
 
-mkdir results/MREC_MAE_80
-python ./gazerr-runner.py data/validation_80_MAE.csv 400 1000 40,40 340,290 results/MREC_MAE_80
+mkdir results/MREC_MErr_80
[…]

### 2022-10-20 scripts/RUN_PRECISE_BIAS_EXPERIMENTS.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+mkdir results/MREC_MErr_Biased_Precise_20
+python ./gazerr-runner.py data/validation_20_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_20
+
+mkdir results/MREC_MErr_Biased_Precise_30
+python ./gazerr-runner.py data/validation_30_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_30
+
+mkdir results/MREC_MErr_Biased_Precise_40
+python ./gazerr-runner.py data/validation_40_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_40
+
+mkdir results/MREC_MErr_Biased_Precise_50
+python ./gazerr-runner.py data/validation_50_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_50
+
+mkdir results/MREC_MErr_Biased_Precise_60
+python ./gazerr-runner.py data/validation_60_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_60
+
+mkdir results/MREC_MErr_Biased_Precise_70
+python ./gazerr-runner.py data/validation_70_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_70
+
+mkdir results/MREC_MErr_Biased_Precise_80
+python ./gazerr-runner.py data/validation_80_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_80
+
+mkdir results/MREC_MErr_Biased_Precise_90
+python ./gazerr-runner.py data/validation_90_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_90
+
+mkdir results/MREC_MErr_Biased_Precise_100
+python ./gazerr-runner.py data/validation_100_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_100
+
+mkdir results/MREC_MErr_Biased_Precise_110
+python ./gazerr-runner.py data/validation_110_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_110
[…]

### 2022-10-20 scripts/generate_datasets.py

@@ -21,7 +21,7 @@ def add_noise(x,y,maerr, bias=False, precise=False):
     y_err = random.uniform(-part*factor,part*factor)
     if bias:
         if precise:
-            y_err = x_err + random.uniform(-0.35*x_err, 0.35*x_err)
+            y_err = x_err + random.uniform(0, 0.3*x_err)
         x_err = -abs(x_err)
         y_err = -abs(y_err)
     return x+x_err,y+y_err
@@ -38,7 +38,7 @@ for mae in maes:
     results3 = pd.DataFrame(columns=header)
     filename = "data/validation_%s_MErr.csv"%str(mae)
     filename2 = "data/validation_%s_MErr_Biased.csv"%str(mae)
-    filename2 = "data/validation_%s_MErr_Biased_Precise.csv"%str(mae)
+    filename3 = "data/validation_%s_MErr_Biased_Precise.csv"%str(mae)
     for n in range(N):
         my_x = random.randrange(0, device_width, 20)
         my_y = random.randrange(0, device_height, 20)
@@ -58,7 +58,7 @@ for mae in maes:
               "gaze_y":gaze_y
            }, ignore_index=True
         )
-        gaze_x, gaze_y = add_noise(my_x,my_y,mae,bias=True,True)
+        gaze_x, gaze_y = add_noise(my_x,my_y,mae,bias=True,precise=True)
         results3 = results3.append({
               "target_x":my_x,
               "target_y":my_y,

### 2022-10-20 scripts/plot_expected_errors.py

@@ -3,8 +3,10 @@ import numpy as np
 from scipy.stats import expon
 import matplotlib.pyplot as plt
 
-file_prefix = "results/MREC_MAE_"
-file_prefix2 = "results/MREC_MAE_Biased_"
+file_prefix = "results/MREC_MErr_"
+file_prefix2 = "results/MREC_MErr_Biased_"
+file_prefix2 = "results/MREC_MErr_Biased_"
+file_prefix3 = "results/MREC_MErr_Biased_Precise_"
 filename = "/expected_values.csv"
 
 errors = [20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
@@ -12,6 +14,8 @@ gaze_err_uni = []
 gaze_err_exp = []
 gaze_err2_uni = []
 gaze_err2_exp = []
+gaze_err3_uni = []
+gaze_err3_exp = []
 
 # This creates the weighting for measurements
 x = expon.rvs(scale=30, size=10000)
@@ -23,10 +27,13 @@ dist_uni = [1/len(vallist) for x in vallist]
 for err in errors:
     file = file_prefix + str(err) + filename
     file2 = file_prefix2 + str(err) + filename
+    file3 = file_prefix3 + str(err) + filename
     df = pd.read_csv(file)
     df2 = pd.read_csv(file2)
+    df3 = pd.read_csv(file3)
     df['error'] = abs(df['Measured']-df['Expected'])
     df2['error'] = abs(df2['Measured']-df2['Expected'])
+    df3['error'] = abs(df3['Measured']-df3['Expected'])
     df['uni'] = dist_uni
     df['exp'] = dist_exp
     df['err_uni'] = df['error']*df['uni'] 
@@ -39,28 +46,44 @@ for err in errors:
     df2['err_exp'] = df2['error']*df2['exp'] 
     gaze_err2_uni.append(df2['err_uni'].sum())
     gaze_err2_exp.append(df2['err_exp'].sum()) 
+    df3['uni'] = dist_uni
+    df3['exp'] = dist_exp
+    df3['err_uni'] = df3['error']*df3['uni'] 
+    df3['err_exp'] = df3['error']*df3['exp'] 
+    gaze_err3_uni.append(df3['err_uni'].sum())
+    gaze_err3_exp.append(df3['err_exp'].sum()) 
 
-fig = plt.figure(figsize=(18,8))
+fig = plt.figure(figsize=(22,8))
 #fig = plt.figure(figsize=plt.figaspect(0.5))
 
[…]

### 2022-10-20 scripts/plot_measured_vs_expected.py

@@ -1,23 +1,28 @@
 import pandas as pd
 import matplotlib.pyplot as plt
 
-file = "results/MREC_MAE_70/expected_values.csv"
-file2 = "results/MREC_MAE_Biased_70/expected_values.csv"
+file = "results/MREC_MErr_70/expected_values.csv"
+file2 = "results/MREC_MErr_Biased_70/expected_values.csv"
+file3 = "results/MREC_MErr_Biased_Precise_70/expected_values.csv"
 
 df = pd.read_csv(file)
 df2 = pd.read_csv(file2)
+df3 = pd.read_csv(file3)
 
 measured = df['Measured']
 expected = df['Expected']
 measured2 = df2['Measured']
 expected2 = df2['Expected']
+measured3 = df3['Measured']
+expected3 = df3['Expected']
 
 fig = plt.figure(figsize=(12,6))
 
 ax = fig.add_subplot()
-ax.set_title('Measured Duration Vs Expected Duration')
-ax.plot(measured, expected, c='orchid', label="Uniform Fixation Error")
-ax.plot(measured2, expected2, c='teal', label="Biased Fixation Error", linestyle='dashed')
+ax.set_title('Measured Duration Vs Expected Duration - 70 Pixel Mean Error')
+ax.plot(measured, expected, c='firebrick', label="Uniform Fixation Error")
+ax.plot(measured2, expected2, c='seagreen', label="Biased Fixation Error", linestyle='dashed')
+ax.plot(measured3, expected3, c='darkorchid', label="Precise Biased Fixation Error", linestyle='dotted')
 p1 = max(max(measured), max(expected))
 p2 = min(min(measured), min(expected))
 ax.plot([p1, p2], [p1, p2], c='lightsteelblue', linestyle='dotted', label='Ideal' )
@@ -26,5 +31,4 @@ ax.set_ylabel('Expected Gaze Duration', fontsize=15)
 ax.legend( prop={'size': 15} )
 
 plt.savefig("results/Measured_vs_expected.png")
-
 

### 2022-10-20 scripts/plot_measured_vs_expected.py
Measured Vs Expected Change

@@ -1,3 +1,4 @@
+import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 
@@ -8,6 +9,10 @@ file3 = "results/MREC_MErr_Biased_Precise_70/expected_values.csv"
 df = pd.read_csv(file)
 df2 = pd.read_csv(file2)
 df3 = pd.read_csv(file3)
+
+df['Expected'] = df['Expected'].replace({'0':np.nan, 0:np.nan})
+df2['Expected'] = df2['Expected'].replace({'0':np.nan, 0:np.nan})
+df3['Expected'] = df3['Expected'].replace({'0':np.nan, 0:np.nan})
 
 measured = df['Measured']
 expected = df['Expected']
@@ -26,8 +31,8 @@ ax.plot(measured3, expected3, c='darkorchid', label="Precise Biased Fixation Err
 p1 = max(max(measured), max(expected))
 p2 = min(min(measured), min(expected))
 ax.plot([p1, p2], [p1, p2], c='lightsteelblue', linestyle='dotted', label='Ideal' )
-ax.set_xlabel('Measured Gaze Duration', fontsize=15)
-ax.set_ylabel('Expected Gaze Duration', fontsize=15)
+ax.set_xlabel('Measured Gaze Duration (ms)', fontsize=15)
+ax.set_ylabel('Expected Gaze Duration (ms)', fontsize=15)
 ax.legend( prop={'size': 15} )
 
 plt.savefig("results/Measured_vs_expected.png")

### 2022-10-20 scripts/RUN_BIAS_EXPERIMENTS.sh
Fix calibration file naming issue

@@ -1,49 +1,49 @@
 #!/bin/bash
 
+mkdir results/MREC_MErr_Biased_70
+python ./gazerr-runner.py data/validation_70_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_70
 
 mkdir results/MREC_MErr_Biased_20
-python ./gazerr-runner.py data/validation_20_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_20
+python ./gazerr-runner.py data/validation_20_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_20
 
 mkdir results/MREC_MErr_Biased_30
-python ./gazerr-runner.py data/validation_30_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_30
+python ./gazerr-runner.py data/validation_30_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_30
 
 mkdir results/MREC_MErr_Biased_40
-python ./gazerr-runner.py data/validation_40_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_40
+python ./gazerr-runner.py data/validation_40_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_40
 
 mkdir results/MREC_MErr_Biased_50
-python ./gazerr-runner.py data/validation_50_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_50
+python ./gazerr-runner.py data/validation_50_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_50
 
 mkdir results/MREC_MErr_Biased_60
-python ./gazerr-runner.py data/validation_60_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_60
+python ./gazerr-runner.py data/validation_60_MErr_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_60
 
-mkdir results/MREC_MErr_Biased_70
-python ./gazerr-runner.py data/validation_70_MAE_Biased.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_70
 
 mkdir results/MREC_MErr_Biased_80
[…]

### 2022-10-20 scripts/RUN_PRECISE_BIAS_EXPERIMENTS.sh

@@ -1,48 +1,49 @@
 #!/bin/bash
 
+mkdir results/MREC_MErr_Biased_Precise_70
+python ./gazerr-runner.py data/validation_70_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_70
+
 mkdir results/MREC_MErr_Biased_Precise_20
-python ./gazerr-runner.py data/validation_20_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_20
+python ./gazerr-runner.py data/validation_20_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_20
 
 mkdir results/MREC_MErr_Biased_Precise_30
-python ./gazerr-runner.py data/validation_30_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_30
+python ./gazerr-runner.py data/validation_30_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_30
 
 mkdir results/MREC_MErr_Biased_Precise_40
-python ./gazerr-runner.py data/validation_40_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_40
+python ./gazerr-runner.py data/validation_40_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_40
 
 mkdir results/MREC_MErr_Biased_Precise_50
-python ./gazerr-runner.py data/validation_50_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_50
+python ./gazerr-runner.py data/validation_50_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_50
 
 mkdir results/MREC_MErr_Biased_Precise_60
-python ./gazerr-runner.py data/validation_60_MAE_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_60
+python ./gazerr-runner.py data/validation_60_MErr_Biased_Precise.csv 400 1000 40,40 340,290 results/MREC_MErr_Biased_Precise_60
 
-mkdir results/MREC_MErr_Biased_Precise_70
[…]

### 2022-10-20 paper/paper.tex
Starting paper modification for precision experiment

@@
 
-The source repository for the gazerr application also contains a series of scripts for generating data
-sets and running the experiments outlined in the next section. Full details available in the package
-README file.
+The source repository for the \textit{gazerr} application also contains a series of 
+scripts for generating data sets and running the experiments outlined in the next 
+section. Full details available in the package \textit{README} file.
 
@@
 
-We apply the gazerr application to investigate the relationship between measured gaze duration
-and expected real duration under a variety of changing scenarios. These scenarios are variations
-in the underlying error indicated by the eye tracking calibration file and length of the session
-in which the user could be looking at the area of interest.
+We apply the \textit{gazerr} application to investigate the relationship between 
+measured gaze duration and expected true duration under a variety of changing scenarios. 
+These scenarios are variations in the underlying error indicated by the eye tracking 
+calibration file and length of the session in which the user could be looking at 
+the area of interest.
 
-For the sake of our simulations we use a simulated device with viewport pixel 
-dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium rectangle ad unit (MREC)
-that is located toward the top of the screen. We iterate over a range of potential mean error
-values (equivalent to varying the accuracy), determined to by the mean Euclidean distance 
+For the sake of our experiments we use a simulated device with viewport pixel 
+dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium 
+rectangle ad unit (MREC) that is located toward the top of the screen. 
+We iterate over a range of potential mean error values 
+(equivalent to varying the accuracy), determined to by the mean Euclidean distance 
 between the measured fixation point and actual fixation points.
-For each of these mean error values we generate a synthetic eye tracking calibration file. We produce
-two versions of this calibration data, one in which the error is centered around the true target,
-and a second biased version in which the error tends toward the upper left of the true position. We 
-include the biased calibration data as it is consistent with our observations of real eye tracking data
-and we explore the extent to which this bias affects gaze duration measurements.
+For each of these mean error values we generate a synthetic eye tracking calibration 
+file. We produce three versions of this calibration data, one in which the error 
+is centered around the true target, a second biased version in which the error tends 
+toward the upper left of the true position. In the third version the error is again
+biased toward the upper left corner of the screen, but it has higher precision, 
+meaning that the measurements will cluster closely together with each other around
+the mean error position.
 
-We feed each of these synthetic calibration files into gazerr to calculate posterior probability
-distributions over true gaze duration for the set of measured gaze durations. These posterior
-distributions are converted into expected values of true gaze duration for each measured duration.
-We repeat the above process for multiple session lengths to investigate the impact of increasing
-session length on the expected gaze duration for a fixed set of measurements.
+We include these biased calibration dataset as they are consistent with both our 
+observations of real eye tracking data and what is found in the literature\cite{}.
+We use this synthetic data creation in order to explore the extent to which different
+patterns in error, bias and precision affects gaze duration measurements.
+
+We feed each of these synthetic calibration files into the \textit{gazerr} application
+to calculate the posterior probability distributions over true gaze duration for 
+the set of measured gaze durations. These posterior distributions are converted 
+into expected values of true gaze duration for each measured duration.
+
+In these experiments we use both a fixed session length and fixed area of interest
+size and position, so that we can focus specifically on the impacts of mean
+error, bias and precision on the quality of gaze duration measurment.
 
@@
 The first experimental result is the expected true duration against the measured duration
-for a single eye tracking application with a mean fixation error of 50 pixels. These
+for a single eye tracking application with a mean fixation error of 70 pixels. These
 results are shown in Figure \ref{fig:measured_vs_expected}.
@@
 \includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
-\caption{Measured versus expected gaze duration. MREC in 10s Session with 50px Mean Error.}
+\caption{Measured Versus Expected Gaze Duration. \\\hspace{\textwidth} 
+MREC ad appearing on screen for 1s with 50px Mean Error in Fixation}
 \label{fig:measured_vs_expected}
@@
 multiple variations of the simulation configuration. At lower meaurement values there is a 
-tendency for the measurement to be an accurate estimation, while at higher values it tends
-toward an under-estimation of the true gaze duration. Surprisingly the results using the
-biased calibration data seem to mitigate this effect somewhat, but the overall pattern 
-remains consistent.
+tendency for the measurement to be an accurate estimation of true gaze duraiton, while 
+at higher values it tends toward an under-estimation of the true gaze duration. 
+Surprisingly the results using the biased calibration data seem to mitigate this effect 
+somewhat, but the overall pattern remains consistent.
 
-We next examined the way that exected error in the gaze duration estimate tracks with expected 
-error in the gaze fixation points. This involved calculating the mean absolute error in gaze
-duration for each of the mean fixation error validation files. We performed this calculation
-using two methods. The first method, called uniform, assumes that all measured values are 
-equally likely. The second method, called smoothed, assumes that low measurements are much
-more likely than higher measurement, and that this drop in likelihood follows an approximately
-exponential distribution.
+We next examined the way that exected error in the gaze duration estimate tracks with 
+expected error in the gaze fixation points. This involved calculating the mean absolute
+error in gaze duration for each of the mean fixation error validation files. 
+We performed this calculation using two methods. The first method, labelled uniform, 
+assumes that all measured values are equally likely. The second method, labelled exponential,
+assumes that low measurements are much more likely than higher measurement, and that 
+this drop in likelihood follows an approximately exponential distribution.
 
@@
 As shown in Figure \ref{fig:fixation_vs_duration} we see that the error in gaze duration
-grows with error in fixation points, as would be expected. However, when we use the exponential
-weighting to calculate the expected error in duration the gaze duration error grows much slower
-than in the uniform case. This suggests that the observed tendency toward smaller measurements
-in area of interest studies means lower expected error overall. This pattern is true for the 
-centrally distributed error (A) as well as the biased calibration error (B).
+grows with error in fixation points, as would be expected. However, when we use the 
+exponential weighting to calculate the expected error in duration the gaze duration 
+error grows much slower than in the uniform case. This suggests that the observed 
+tendency toward smaller measurements in advertising area of interest studies means 
+lower expected error overall. This pattern is true for the centrally distributed 
+error (A) as well as the biased calibration error (B) and precise biased calibration
+error (C) scenarios.
 
@@
 
-The left hand plot shows the complete error surface in which you will observe large spikes in
+The left hand plot shows the complete error surface in which you will observe large 
+spikes in
 error in the corner corresponding to both high mean error for the fixation models and 

### 2023-03-01 paper/paper.tex
Uncommited paper changes

@@
 
-\subsection{Experiments}
+\section{Experiments}
 
 We apply the \textit{gazerr} application to investigate the relationship between 
-measured gaze duration and expected true duration under a variety of changing scenarios. 
-These scenarios are variations in the underlying error indicated by the eye tracking 
+measured gaze duration and expected true duration under a variety of scenarios. 
+Each scenario involves a variation in the underlying error indicated by the eye tracking 
 calibration file and length of the session in which the user could be looking at 
@@
 
-We include these biased calibration dataset as they are consistent with both our 
+We include these biased calibration datasets as they are consistent with both our 
 observations of real eye tracking data and what is found in the literature\cite{}.
@@
 \includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
-\caption{Measured Versus Expected Gaze Duration. \\\hspace{\textwidth} 
-MREC ad appearing on screen for 1s with 50px Mean Error in Fixation}
+\caption{Measured Gaze Duration Versus Expected Gaze Duration. \\\hspace{\textwidth} 
+MREC ad appearing on screen for 1s with 70px Mean Error in Fixation}
 \label{fig:measured_vs_expected}
@@
 
-The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we see under
-multiple variations of the simulation configuration. At lower meaurement values there is a 
-tendency for the measurement to be an accurate estimation of true gaze duraiton, while 
+The result shown in Figure \ref{fig:measured_vs_expected} is indicative of what we 
+see under multiple variations of the simulation configuration. 
+At lower meaurement values there is a tendency for the measurement to be an 
+accurate estimation of true gaze duration, while 
 at higher values it tends toward an under-estimation of the true gaze duration.
-Surprisingly the results using the biased calibration data seem to mitigate this effect 
-somewhat, but the overall pattern remains consistent.
+This problem is worse for the model with the biased error, but it is mitigated
+when the biased error comes from a model with higher precision. This is potentially
+due to the fact that the high prescion scenario has lower tendency for gaze to be 
+pushed off screen.
 
@@
 
+In addition, the additional error in gaze duration we see when the error is biased
+in a particular direction (shown in B) is partly mitigated by a fixation model that
+is more precise (shown in C). This suggests that a precise fixation sampling process
+is able to partly overcome the false positive and false negatives of an error prone gaze
+fixation model.
+
 In our final experiment we look at the effect of both the size of the measured 
 gaze duration and the mean error of fixation on the expected error in gaze duration.
-We display this as a three dimensional expected error surface over these two key dimensions
-in Figure \ref{fig:error_surface}
+We conduct this experiment using only the unbiased synthetic data in order to understand
+how error centred on true fixation points relates to error in gaze duration across the
+range of measurement. We display these results as a three dimensional expected error 
+surface over these two key dimensions in Figure \ref{fig:error_surface}
 
@@
 simulating the distribution of measured duration for each possible true duration, 
-and then inverting it with Bayes rule into a distribution over true duration for a given measurement.
+and then inverting it with Bayes rule into a distribution over true duration for a 
+given measurement.
 
-We have released this approach as the open source application \textit{gazerr} and used it to explore
-the relationship between fixation error and duration error. Our results show that the nature
-of the error in gaze duration depends on the size of the measurement. Smaller measurements
-tend to have lower expected error while larger measurements tend toward under prediction. We plotted
-the relationship between expected error and saw that observed tendency toward smaller duration
-measurements (estimated with an psuedo-exponential distribution) delivers a slower growing
+We have released this approach as the open source application \textit{gazerr} and 
+used it to explore the relationship between fixation error and duration error, under
+range of realistic error profiles when looking at measuring attention on a standard
+size ad within a mobile screen. Our results show that the nature
+of the error in gaze duration tends to depend on the size of the measurement. 
+Smaller measurements tend to have lower expected error while larger measurements 
+tend toward under prediction. 
+We plotted the relationship between expected error and saw that the observed 
+tendency toward smaller duration measurements 
+(estimated with an psuedo-exponential distribution) delivers a slower growing
 expected duration error with the mean error in fixation points.
 
-Finally, our simulations revealed that for large error in fixation points, certain measurements
-become so unlikely that our Monte Carlo simualtions placed none of the probability density
-in those regions. This strongly suggests that an additional downside to large fixation point error
+Finally, our simulations revealed that for large error in fixation points, 
+certain measurements become so unlikely that our Monte Carlo simulations placed 
+none of the probability density in those regions. 
+This strongly suggests that an additional downside to large fixation point error
 is a reduced range of measurement for area of interest studies.

### 2023-03-01 scripts/plot_expected_errors.py

@@ -53,7 +53,7 @@ for err in errors:
     gaze_err3_uni.append(df3['err_uni'].sum())
     gaze_err3_exp.append(df3['err_exp'].sum()) 
 
-fig = plt.figure(figsize=(22,8))
+fig = plt.figure(figsize=(18,8))
 #fig = plt.figure(figsize=plt.figaspect(0.5))
 
 # =============
@@ -61,7 +61,7 @@ fig = plt.figure(figsize=(22,8))
 # =============
 # set up the axes for the first plot
 ax = fig.add_subplot(1, 3, 1)
-ax.set_title('A) Expected Error - Gaze Fixation Vs Duration')
+ax.set_title('A) Gaze Fixation Vs Duration Error')
 ax.plot(errors, gaze_err_uni, c='firebrick', label="Uniform Measurement Distribution")
 ax.plot(errors, gaze_err_exp, c='salmon', linestyle='dashed', label="Exponential Measurement Distribution")
 ax.set_xlabel('Gaze Fixation μ Error (px)', fontsize=12)
@@ -71,7 +71,7 @@ ax.set_ylim([0, 500])
 ax.legend( prop={'size': 10} )
 
 ax = fig.add_subplot(1, 3, 2)
-ax.set_title('B) Expected Error - Biased Gaze Fixation Vs Duration')
+ax.set_title('B) Biased Gaze Fixation Vs Duration Error')
 ax.plot(errors, gaze_err2_uni, c='seagreen', label="Uniform Measurement Distribution")
 ax.plot(errors, gaze_err2_exp, c='lightseagreen', linestyle='dashed', label="Exponential Measurement Distribution")
 ax.set_xlabel('Gaze Fixation μ Error (px)', fontsize=12)
@@ -81,7 +81,7 @@ ax.set_ylim([0, 500])
 ax.legend( prop={'size': 10} )
 
 ax = fig.add_subplot(1, 3, 3)
-ax.set_title('C) Expected Error - Biased Precise Gaze Fixation Vs Duration')
+ax.set_title('C) Precise Biased Fixation Vs Duration Error')
 ax.plot(errors, gaze_err3_uni, c='darkorchid', label="Uniform Measurement Distribution")
 ax.plot(errors, gaze_err3_exp, c='orchid', linestyle='dashed', label="Exponential Measurement Distribution")
 ax.set_xlabel('Gaze Fixation μ Error (px)', fontsize=12)

### 2023-03-01 paper/paper.tex
Paper update

@@
 for a given true duration. Secondly, inverting this into the distribution of true
-gaze durtions for a given measurement. Note, that the algorithm requires that the area of
+gaze durations for a given measurement. Note, that the algorithm requires that the area of
 interest be defined as a fixed bounding box and that the distributions are estimated for
@@
 $generateMeasurement(p,l,C)$, takes the path, target location and the eye tracking calibration 
-data and produces a sample of measured attention time. It does this by using the calibration data to sample
-fixation errors and apply them to the path data. The noisy path data is then used to determine
+data and produces a sample of measured attention time.
+This is achieved by using the calibration data to sample
+fixation errors in the proximity of each location within the path and applying them to the path data. 
+The noisy path data is then used to determine
 the estimated gaze duration by looking at the number of fixation points that intersect with the target
@@
 values of $d$ allows us to populate a two dimensional array where the first dimension is the
-true duration and the second the estimated duration. Note that our probability distribution is discrete
+true duration and the second the estimated duration.
+
+Note that our probability distribution is discrete
 because we restrict ourselves to sampling over a set of fixed time period increments ($i$) 
-between zero and the session length.
+between zero and the session length. This restriction is realistic for most machine learning
+eye tracking systems as they output fixation points at fixed intervals of time.
 
@@
 or it can deployed as a command line application. 
-The source code is available on GitHub and the python package can be installed from PyPi.
+The source code is available on GitHub\footnote{https://github.com/playground-xyz/gazerr}
+and the python package can be installed from PyPi.\footnote{https://pypi.org/project/gazerr/}
 
@@
 scripts for generating data sets and running the experiments outlined in the next 
-section. Full details available in the package \textit{README} file.
+section. Full details for the data generation process are available in the 
+source code README file.
 
@@
 We include these biased calibration datasets as they are consistent with both our 
-observations of real eye tracking data and what is found in the literature\cite{}.
+observations of real eye tracking data and what is found in the literature\cite{Holmqvist2022}.
 We use this synthetic data creation in order to explore the extent to which different

### 2023-03-01 acm_version/paper.tex
ACM Version of Paper

--- 2023-03-01 paper/paper.tex
@@
-\documentclass[12pt,a4paper]{article}
+%%
+%% This is file `sample-sigconf.tex',
+%% generated with the docstrip utility.
+%%  
+%% The original source files were: 
+%%
+%% samples.dtx  (with options: `sigconf')
+%%
+%% IMPORTANT NOTICE:
+%% 
+%% For the copyright see the source file.
+%% 
+%% Any modified versions of this file must be renamed
+%% with new filenames distinct from sample-sigconf.tex.
+%% 
+%% For distribution of the original source see the terms
+%% for copying and modification in the file samples.dtx.
+%% 
+%% This generated file may be distributed as long as the
+%% original source files, as listed above, are part of the
+%% same distribution. (The sources need not necessarily be
+%% in the same archive or directory.)
+%%
+%%
+%% Commands for TeXCount
+%TC:macro \cite [option:text,text]
+%TC:macro \citep [option:text,text]
+%TC:macro \citet [option:text,text]
+%TC:envir table 0 1
+%TC:envir table* 0 1
+%TC:envir tabular [ignore] word
+%TC:envir displaymath 0 word
+%TC:envir math 0 word
+%TC:envir comment 0 0
+%%
+%%
+%% The first command in your LaTeX source must be the \documentclass command.
+\documentclass[sigconf]{acmart}
+
+%% Rights management information.  This information is sent to you
+%% when you complete the rights form.  These commands have SAMPLE
+%% values in them; it is your responsibility as an author to replace
+%% the commands and values with those provided to you when you
+%% complete the rights form.
+\setcopyright{acmcopyright}
+\copyrightyear{2023}
+\acmYear{2023}
+\acmDOI{XXXXXXX.XXXXXXX}
+
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[MLHCI '23]{}{March 24--26,2023}{Singapore}
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/23/05}
+
+%%
+%% Submission ID.
+%% Use this when submitting an article to a sponsored event. You'll
+%% receive a unique submission ID from the organizers
+%% of the event, and this ID should be used as the parameter to this command.
+\acmSubmissionID{VP5015}
+
+%%
+%% You can then either use BibTeX with the ACM-Reference-Format style,
+%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
+%% support for advanced citation of software artefact from the
+%% biblatex-software package, also separately available on CTAN.
+%%
+%% Look at the sample-*-biblatex.tex files for templates showcasing
+%% the biblatex styles.
+%%
+%%
+%% The majority of ACM publications use numbered citations and
+%% references.  The command \citestyle{authoryear} switches to the
+%% "author year" style.
+%%
+%% If you are preparing content for an event
+%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
+%% citations and references.
+%% Uncommenting
+%% the next command will enable that style.
+%%\citestyle{acmauthoryear}
+
 \usepackage{amsfonts}
@@
 \usepackage{hyperref}
-\usepackage[margin=2.9cm]{geometry}
-\usepackage[outdir=./images/]{epstopdf}
-\epstopdfsetup{outdir=./images/}
 \usepackage{algorithm}
@@
 
-\author{
-John Hawkins \\ email \href{mailto:john.hawkins@playgroundxyz.com}{john.hawkins@playgroundxyz.com} \\
+%%
+%% The "author" command and its associated commands are used to define
+%% the authors and their affiliations.
+%% Of note is the shared affiliation of the first two authors, and the
+%% "authornote" and "authornotemark" commands
+%% used to denote shared contribution to the research.
+\author{John Hawkins}
+\email{john.hawkins@playgroundxyz.com}
+\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{Playground XYZ}
+  \streetaddress{Level 13/50 Carrington St}
+  \city{Sydney}
+  \state{NSW}
+  \country{Australian}
+  \postcode{2000}
 }
+\renewcommand{\shortauthors}{Hawkins}
 
-\maketitle
-
-\section{Abstract}
-
+\begin{abstract}
 Eye tracking technologies produce a series of timestamped gaze fixation points 
@@
 and the comparative length of the viewing session.
+\end{abstract}
+
+%%
+%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+%% Please copy and paste the code instead of the example below.
+%%
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10002951.10003260.10003272</concept_id>
+       <concept_desc>Information systems~Online advertising</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10002951.10003260.10003261.10003267</concept_id>
+       <concept_desc>Information systems~Content ranking</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10002951.10003260.10003277.10003281</concept_id>
+       <concept_desc>Information systems~Traffic analysis</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Information systems~Online advertising}
+\ccsdesc[500]{Information systems~Content ranking}
+\ccsdesc[500]{Information systems~Traffic analysis}
+
+\keywords{Eye Tracking, Attention Measurement, Digital Advertising, Error Estimation}
+
+\maketitle
 
@@
 
-The source repository for the \textit{gazerr} application also contains a series of 
-scripts for generating data sets and running the experiments outlined in the next 
-section. Full details for the data generation process are available in the 
-source code README file.
+The source repository for the \textit{gazerr} application also contains the complete
+set of scripts for re-creating the data sets and running the experiments outlined in the next 
+section. Please refer to the README file for the sequence of steps required for replication
+of all results.
 
@@
 We apply the \textit{gazerr} application to investigate the relationship between 
-measured gaze duration and expected true duration under a variety of scenarios. 
+measured gaze duration and expected true duration of fixation under a variety of 
+simulated scenarios. 
 Each scenario involves a variation in the underlying error indicated by the eye tracking 
@@
 
-For the sake of our experiments we use a simulated device with viewport pixel 
-dimensions of 350 by 627. Our area of interest (AOI) is a fixed position medium 
-rectangle ad unit (MREC) that is located toward the top of the screen. 
-We iterate over a range of potential mean error values 
-(equivalent to varying the accuracy), determined to by the mean Euclidean distance 
-between the measured fixation point and actual fixation points.
-For each of these mean error values we generate a synthetic eye tracking calibration 
-file. We produce three versions of this calibration data, one in which the error 
-is centered around the true target, a second biased version in which the error tends 
-toward the upper left of the true position. In the third version the error is again
-biased toward the upper left corner of the screen, but it has higher precision, 
-meaning that the measurements will cluster closely together with each other around
-the mean error position.
+For the sake of our experiments we use a simulated device, approximately the size of a
+small smart-phone, with viewport pixel dimensions of 350 by 627. 
+Our area of interest (AOI) is a fixed position 
+rectangle that resembles an iAB standard medium rectangle ad unit (MREC). We position the
+simulated ad toward the top of the screen. 
+We iterate over a range of potential mean error values from 20 to 160 pixels 
+(equivalent to varying the accuracy of the eye tracking models). These mean error
+values determine the average Euclidean distance 
+between a measured fixation point and the true fixation point.
 
-We include these biased calibration datasets as they are consistent with both our 
+For each of these mean error values we can generate synthetic eye tracking calibration 
+files containing a sample of true fixation points and measured fixation points that are
+coherent with the simulation parameters. 
+We produce three versions of this calibration data to simulate and experiment with
+multiple real-world scenarios:
+
+\begin{itemize}
+    \item {\texttt{Unbiased}}: Error distributed uniformly around true fixation point. 
+    \item {\texttt{Biased}}: Error tending toward upper left of true fixation.
+    \item {\texttt{Biased Precise}}: Error tending toward upper left but clustered.
+\end{itemize}
+
+These three scenarios are listed above. The unbiased calibration data is
+distributed evenly and centered around each of the true target positions in the callibration
+file. The biased data imposes a tendency on the error to be pushed toward the upper left 
+of the screen away from the true fixation position. The biased precise data is again forced
+toward the upper left corner of the screen, but constrained so that measurements cluster
+together, meaning that although measurements have the same mean error that are more tightly
+clustered with each other around the mean error position.
+
+We include the two biased calibration datasets as they are consistent with both our 
 observations of real eye tracking data and what is found in the literature\cite{Holmqvist2022}.
 We use this synthetic data creation in order to explore the extent to which different
-patterns in error, bias and precision affects gaze duration measurements.
+patterns in error, bias and precision will affects gaze duration measurements.
 
@@
 
-In these experiments we use both a fixed session length and fixed area of interest
-size and position, so that we can focus specifically on the impacts of mean
-error, bias and precision on the quality of gaze duration measurment.
+In these experiments we use both a fixed session length of one thousand milliseconds
+and a fixed position for the area of interest. These restrictions allow us to focus 
+on the impacts of mean error, bias and precision on the quality of gaze duration measurment.
 
@@
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
 \includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
@@
 \label{fig:measured_vs_expected}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
 \centering
@@
 \label{fig:fixation_vs_duration}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:error_surface}
-\end{figure}
+\end{figure*}
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{refs}
@@
 \end{document}
+\endinput
+%%
+%% End of file `sample-sigconf.tex'.

### 2023-03-01 acm_version/paper.tex
ACM Categories

@@
 %%
+
 \begin{CCSXML}
 <ccs2012>
+   <concept>
+       <concept_id>10002950.10003648.10003671</concept_id>
+       <concept_desc>Mathematics of computing~Probabilistic algorithms</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10003120.10003121</concept_id>
+       <concept_desc>Human-centered computing~Human computer interaction (HCI)</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
    <concept>
@@
        </concept>
-   <concept>
-       <concept_id>10002951.10003260.10003261.10003267</concept_id>
-       <concept_desc>Information systems~Content ranking</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10002951.10003260.10003277.10003281</concept_id>
-       <concept_desc>Information systems~Traffic analysis</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
  </ccs2012>
@@
 
+\ccsdesc[500]{Mathematics of computing~Probabilistic algorithms}
+\ccsdesc[500]{Human-centered computing~Human computer interaction (HCI)}
 \ccsdesc[500]{Information systems~Online advertising}
-\ccsdesc[500]{Information systems~Content ranking}
-\ccsdesc[500]{Information systems~Traffic analysis}
 

### 2023-03-24 acm_version/paper.tex
Paper update for conference

@@
 \label{eq:p_of_d}
-P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ P(\hat{d},l,s)  }
+P(d|\hat{d},l,s) =  \frac{ P(\hat{d}|l,s,d) P(d|l,s) }{ P(\hat{d}|l,s)  }
 \end{equation}
 
-We use a uniform prior for $P(d)$, meaning in the absence of additional information all
+We use a uniform prior for $P(d|l,s)$, meaning in the absence of additional information all
 gaze durations less than the session length are equally likely. As our distributions are
-discrete estimations of an underlying continuous distribution the value of $P(d)$ is
-equal to $1/(1+floor(i*s))$. We can estimate the value
-of the denominator $P(\hat{d},l,s)$ by iterating over all values of $d$ and summing the product
-of $ P(\hat{d},l,s|d) \dot P(d)$. This fully explicated form is shown in 
-Equation \ref{eq:p_of_d_full}.
+discrete estimations of an underlying continuous distribution the value of each discrete
+value in $P(d|l,s)$ is
+equal to $1/(1+floor(s/i))$. We can estimate the value
+of the denominator $P(\hat{d}|l,s)$ by iterating over all possible values of $\delta \in D$ 
+and summing the product of $P(\hat{d}|l,s,\delta)$ and $P(\delta|l,s)$. 
+This fully explicated form is shown in Equation \ref{eq:p_of_d_full}.
 
@@
 \label{eq:p_of_d_full}
-P(d|\hat{d},l,s) =  \frac{ P(\hat{d},l,s|d) \dot P(d) }{ \sum_{\delta \in D} P(\hat{d},l,s|\delta) \dot P(\delta)  }
+P(d|\hat{d},l,s) =  \frac{ P(\hat{d}|l,s,d) P(d|l,s) }{ \sum_{\delta \in D} P(\hat{d}|l,s,\delta) P(\delta|l,s) }
 \end{equation}
@@
 \begin{figure*}[h!]
-\includegraphics[scale=0.5]{../results/Measured_vs_expected.png}
+\includegraphics[scale=0.5]{results/Measured_vs_expected.png}
 \caption{Measured Gaze Duration Versus Expected Gaze Duration. \\\hspace{\textwidth} 
@@
 \centering
-\includegraphics[scale=0.4]{../results/Fixation_vs_duration_error.png}
+\includegraphics[scale=0.4]{results/Fixation_vs_duration_error.png}
 \caption{Fixation error versus gaze duration error.}
@@
 \centering
-\includegraphics[scale=0.4]{../results/Error_surface.png}
+\includegraphics[scale=0.4]{results/Error_surface.png}
 \caption{Mean gaze duration error over measured error and fixation MAE.}
